# Data Preparation
This chapter documents our decisions surrounding how to prepare the data for analysis. These decisions include details on loading the necessary R packages, loading the data in R, coding or recoding of variables. Also discussed is how we deal with missing data.

## Preliminary work

**Load R packages**. Some R users load all their packages up front. Others, perhaps for instructional reasons, load them just before calling functions, perhaps for instructional reasons. I can see merit in listing all the packages for properly citing them. Here is a code chunk for loading all the packages up front:

```{r include=TRUE, message=FALSE}
# LOAD PACKAGES
library(Hmisc) # for variable and value labels
library(naniar) # for missing data
library(corrplot) # for correlation plot
library(knitr) # how RMarkdown compiles code into documents
```

**Load data**. The data are contained in an Excel file, simply named "data.xlsx", in the "data" folder of this project. The important thing is to keep all the data in the same folder within the larger "SHAS Validity" project folder.

**Select items for analysis**. Which items should we include for analysis? The first 13 columns in the original data file report data from several screening items that were placed at the beginning of the questionnaire. But because they are constant and not variable, they don't seem to have a place in scale development. Therefore it seems the important columns range from columns 14 to 97. The first set of items measure the **External Events** scale. These 52 items range from columns 14 to 35. The second set of items measure the **Internal States** scale, and these 14 items range from columns 66 to 79. The last set of items capture **demographic** variables, and these range from columns 80 to 97.

```{r}
# LOAD ORIGINAL DATA
library(readxl) # for reading the source Excel file
shas <- read_excel("data/data.xlsx") # load Excel file

# SELECT DATASETS FOR ANALYSIS
all <- shas[ ,14:97] # select items, dems from columns 14-79
all <- subset(all, select = -c(Ext_Avg:Global_Avg)) # remove columns for now
items <- all[ ,1:66]
ext_items <- items[ ,1:52]
int_items <- items[ ,52:66]

# SAVE DATA FILES
write.csv(all,"data/all.csv", row.names=FALSE)
write.csv(items,"data/items.csv", row.names=FALSE)
write.csv(ext_items,"data/ext_items.csv", row.names=FALSE)
write.csv(int_items,"data/int_items.csv", row.names=FALSE)
```

## Missing data
Missing data can be an important problem. The problem has two parts. The first is how many data are missing. Obviously the more data are missing, the larger the barrier this problem poses to analysis. The second is the mechanism of missing data. Are the missing data randomly, or is there a pattern to the missing data - like a problematic item, or a distinct group of examinees who chose not to respond? Figure \@ref(fig:vim-plots) presents two plots from the VIM package for missing data [@R-VIM]. 

```{r vim-plots, fig.cap='Missing Data', message=FALSE, warning=FALSE}
# VISUALIZE AND QUANTIFY MISSING DATA
library(VIM)
a <- VIM::aggr(all, plot = FALSE)
plot(a, numbers = TRUE, prop = FALSE)
```

The first plot in Figure \@ref(fig:vim-plots) is a frequency distribution of the number of missing data entries for each item. This plot reveals very few missing data for the first set of External Events items, but nearly 200 missing entries for the second set of Internal States items and demographics. The raises two questions. The first is: Who were those examinees? We know that they responded to the first 52 items, then apparently quit the survey. Why? Was it just fatigue, or is there more to know about them? The second question is what to do with the information they actually did provide.

The second plot in Figure \@ref(fig:vim-plots) illustrates the different combinations of missing entries. The blue squares represent cases with complete responses, and the red indicates missing data. The rows represent combinations of complete and missing responses. The bottom row of entirely blue squares represents complete data -- examinees who responded to all of the items. The barplot on the right axis indicates that these were the most frequent combinations. The second most frequent combination was examinees who completed all of the External Events items and then quit the survey. Beyond that, there is no apparent order or pattern to the missing data.

```{r eval=FALSE, include=FALSE}
# VISUALIZE AND QUANTIFY MISSING DATA
# library(naniar)
# naniar::vis_miss(all)
```

**Mechanism of missing data**. @McKnightEtAl2007 prescribe several approaches for investigating the mechanisms of missing data. An appropriate first step is to test whether the data are Missing Completely At Random (MCAR) with the use of Little's test @Little1988. This is a chi-square test of the difference between the observed pattern of missing responses and the pattern of responses that would expected if missingness were completely random. We report the results of this test in Table \@ref(tab:mcar).

```{r mcar, message=FALSE, warning=FALSE}
library(naniar)
knitr::kable(mcar_test(all), caption = "Little's Test of Missing Completely At Random (MCAR)")
```

The test is fairly conclusive. The p-value for the chi-square is 1, which means observed pattern of missing data do not diverge significantly from what would be expected if the data were missing completely at random.

**Were examinees triggered?** As mentioned above, a possible explanation of the missing data in the Internal States items is that examinees were triggered. As after responding to the preceding 52 items calling to mind imagery of abusive external events, these examinees were **_done_** with the survey. A possible test of this hypothesis is an examination of the relationship between a total score of responses to the External Events items and a score of total missingness on the Internal States items. The empirical work to test this hypothesis is laid out in the following code chunk. The items measuring external events are summed into a total score such that a higher value indicates more exposure to abusive external events. The items measuring internal states are coded so that 1=missing and 0=all others response, and these items are summed into a score of total missingness.

```{r, message=FALSE, warning=FALSE}

# 1. ISOLATE THE DATA
miss_set <- items

# 2. DUMMY CODE INTERNAL STATES ITEMS
miss_set$Q68[miss_set$Q68 > 0] <- 0
miss_set$Q69[miss_set$Q69 > 0] <- 0
miss_set$Q70[miss_set$Q70 > 0] <- 0
miss_set$Q71[miss_set$Q71 > 0] <- 0
miss_set$Q72[miss_set$Q72 > 0] <- 0
miss_set$Q73[miss_set$Q73 > 0] <- 0
miss_set$Q74[miss_set$Q74 > 0] <- 0
miss_set$Q76[miss_set$Q76 > 0] <- 0
miss_set$Q77[miss_set$Q77 > 0] <- 0
miss_set$Q78[miss_set$Q78 > 0] <- 0
miss_set$Q79[miss_set$Q79 > 0] <- 0
miss_set$Q80[miss_set$Q80 > 0] <- 0
miss_set$Q81[miss_set$Q81 > 0] <- 0
miss_set$Q82[miss_set$Q82 > 0] <- 0
miss_set$Q68[is.na(miss_set$Q68)] <- 1
miss_set$Q69[is.na(miss_set$Q69)] <- 1
miss_set$Q70[is.na(miss_set$Q70)] <- 1
miss_set$Q71[is.na(miss_set$Q71)] <- 1
miss_set$Q72[is.na(miss_set$Q72)] <- 1
miss_set$Q73[is.na(miss_set$Q73)] <- 1
miss_set$Q74[is.na(miss_set$Q74)] <- 1
miss_set$Q76[is.na(miss_set$Q76)] <- 1
miss_set$Q77[is.na(miss_set$Q77)] <- 1
miss_set$Q78[is.na(miss_set$Q78)] <- 1
miss_set$Q79[is.na(miss_set$Q79)] <- 1
miss_set$Q80[is.na(miss_set$Q80)] <- 1
miss_set$Q81[is.na(miss_set$Q81)] <- 1
miss_set$Q82[is.na(miss_set$Q82)] <- 1

# 3. TOTAL SCORES ON EXTERNAL EVENTS ITEMS AND
#    INTERNAL STATES MISSINGNESS
miss_set$total_ext <- rowSums(miss_set[ , c(1:52)])
miss_set$total_int <- rowSums(miss_set[ , c(53:66)])
miss_set2 <- miss_set[ , c(53:68)] # just the total score, missingness
```

```{r trigger, warning=FALSE, message=FALSE, fig.cap='Correlation Matrix: External States and Internal States Missingness'}
# 4. CORRELATIONS
psych::corPlot(miss_set2[,1:16], main=NULL, upper = FALSE)
```

Figure \@ref(fig:trigger) illustrates the correlations among the Internal States items (dichotomized according to their missingness (1=missing, 0=all others)), the total score from the external states (total_ext), and the total score of internal item missingness (total_int). The correlation of total_ext with all the other items is nearly zero, suggesting that more exposure to abusive external states did **_not_** predict nonresponse to the Internal States items. This finding is consistent with the finding from Figure \@ref(tab:mcar) that the missing data appear to missing completely at random.

**Missing data methods**. There are several different ways to handle missing data. A common method is to _delete all cases missing any data_. This is called **listwise** deletion of data. Its primary advantages are simplicity, expedience, and analysis based on complete information. The disadvantages of listwise are (1) disposal of the information offered by those examinees who responded to some items but not others, and (2) diminishment of the dataset, and with it, a reduction in statistical power. The "all" data file of all Likert items and demographics items derived earlier reports data from 3,222 examinees. The following code chunk creates a subset of complete data. This smaller dataset has 1,841 examinees.

```{r listwise, include=TRUE, warning=FALSE, tidy=TRUE}

# LISTWISE DELETION OF MISSING DATA
all_l <- na.omit(all) #  Likert items from columns 14-79 plus demographics post listwise
write.csv(all_l,"data/all_l.csv", row.names=FALSE) # save data

# OTHER POSSIBLE LISTWISE DATASETS

# items <- items[complete.cases(items), ] # create new dataset without missing data
items_l <- na.omit(items) # Likert items from columns 14-79 post listwise
ext_items_l <- na.omit(ext_items) # Likert items for external events post listwise
int_items_l <- na.omit(int_items) # Likert items for internal states post listwise
write.csv(items_l,"data/items_l.csv", row.names=FALSE)
write.csv(ext_items_l,"data/ext_items_l.csv", row.names=FALSE)
write.csv(int_items_l,"data/int_items_l.csv", row.names=FALSE)
```

Analysis can proceed with these two datasets:

1. the "all" dataset of all Likert and demographic items (N=3,222)
2. the "all_l" dataset of 1,841 examinees who reported complete data