# Item Analysis

## Load packages and data

```{r}

# LOAD PACKAGES
library(psych) # for psychometric analysis
library(psychTools) # for psychometric analysis
library(CTT) # for classical test theory stats
library(pastecs) # for item statistics
library(ltm) # for point biserial correlations
library(corrplot) # for correlation plot
library(knitr) # how RMarkdown compiles all the code into documents

# LOAD DATA
# Datasets straight off the Excel file
all         <- read.csv("data/all.csv")
items       <- read.csv("data/items.csv")
ext_items   <- read.csv("data/ext_items.csv")
int_items   <- read.csv("data/int_items.csv")
# Datasets with listwise deletion
all_l       <- read.csv("data/all_l.csv")
items_l     <- read.csv("data/items_l.csv")
ext_items_l <- read.csv("data/ext_items_l.csv")
int_items_l <- read.csv("data/int_items_l.csv")

```

A good way to begin the analysis is to get a feel for our items with descriptive statistics.

```{r}

# DESCRIBE DATA
# options(digits=2)
itm_stats <- describe(all)
knitr::kable(itm_stats, caption = "Items Statistics", booktabs = TRUE, digits=2)

library(pastecs)
item_stats <- pastecs::stat.desc(all) # use the stat.desc() to render data frame of item stats
item_stats <- t(item_stats) # transpose the data frame
item_stats <- data.frame(item_stats)
item_stats[order(item_stats$mean), ] # order by mean
item_stats[order(item_stats$nbr.val), ] # order by N cases

# A POSSIBLE BAR PLOT OF ITEM STATISTICS
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=TRUE, col='red')
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=FALSE, col='red', ylim = c(0,8000),names.arg=labs)

# OUTLIERS
# psych::outlier(items)

# CLASSICAL ITEM ANALYSIS
# CTT::itemAnalysis(items, itemReport=TRUE, NA.Delete=TRUE, rBisML=FALSE, hardFlag, easyFlag, pBisFlag, bisFlag, flagStyle = c("X",""))

# DISTRACTOR ANALYSIS
# CTT::distractorAnalysis

```

### Skew and Kurtosis

The item for gratitude is definitely skewed, based on the skew and kurtosis statistics in the table.  

I will supplement this table with a set of plots to visualize the data. Here I pass the 'table' function into an object for each item. Then I pass these objects into plots of uniform size and labeling.

```{r}

#{r fig.width=10, fig.height=7, fig.cap="Frequencies of Well-Being Items", # fig.align='center'}

# FREQUENCIES
# peacecounts <- table(items$sense.peace)
# wondercounts <- table(items$sense.wonder)
# gratitudecounts <- table(items$sense.gratitude)
# purposecounts <- table(items$sense.purpose)
# labs <- c("1 Never", "2 Seldom", "3 Several times a year", "4 Once or twice a month", "5 At least once a week")
# par(mfrow=c(2,2))
# barplot(peacecounts, main="Deep Sense of Peace", horiz=FALSE, col='red', ylim = c(0,8000),
#         names.arg=labs)
# barplot(wondercounts, main="Deep Sense of Wonder", horiz=FALSE, col='blue', ylim = c(0,8000),
#         names.arg=labs)
# barplot(gratitudecounts, main="Strong Sense of Gratitude", horiz=FALSE, col='purple', ylim = c(0,8000),
#         names.arg=labs)
# barplot(purposecounts, main="Wonder about Meaning and Purpose", horiz=FALSE, col='green', ylim = c(0,8000),
#         names.arg=labs)
```

The striking thing about these data is their negative skew: the vast majority of respondents report frequently feeling peace, wonder, gratitude, and wonder about the meaning of life.  This skew poses some challenges.  One is less variance.  If the vast majority of Catholics past and present report frequent feelings of well-being, there is not much to explain.  The other challenge is that classical statistics are based the assumption that the data are normally distributed.

## Item correlations

```{r}
# CORRELATION MATRIX (LOWER HALF)
# r <- lowerCor(items) # the correlation matrix, rounded to 2 decimals

# HEATMAP OF CORRELATIONS
# corPlot(r) #examine the many options for this function.
# corr.test(items)

# POINT BISERIAL CORRELATIONS
# ltm::biserial.cor(x, y, use = c("all.obs", "complete.obs"), level = 1)

```

## Internal consistency reliability

```{r}

# RELIABILITY
# reliability(items, itemal = TRUE, NA.Delete = TRUE, ml = TRUE)

```

