# Item Analysis
This chapter documents our item analysis. This includes investigation of:

- Item difficulty
- Item discrimination
- Item-total correlations

**Load packages**. This code chunk outlines the R packages we use and the data. There are several helpful packages for classical test theory (CTT).

```{r}
# LOAD PACKAGES
library(psych) # for psychometric analysis
library(psychTools) # for psychometric analysis
library(CTT) # for classical test theory stats
library(pastecs) # for item statistics
library(ltm) # for point-biserial correlations
library(corrplot) # for correlation plot
library(knitr) # how RMarkdown compiles all the code into documents
```

**Load data**. This code chunk outlines the R packages we use and the data. There are several helpful packages for classical test theory (CTT).

```{r}
# LOAD DATA
# Datasets straight off the Excel file
all         <- read.csv("data/all.csv")
# items       <- read.csv("data/items.csv")
# ext_items   <- read.csv("data/ext_items.csv")
# int_items   <- read.csv("data/int_items.csv")
# Datasets with listwise deletion
all_l       <- read.csv("data/all_l.csv")
# items_l     <- read.csv("data/items_l.csv")
# ext_items_l <- read.csv("data/ext_items_l.csv")
# int_items_l <- read.csv("data/int_items_l.csv")
```

**Describe data**. A good way to begin the analysis is to get a feel for our items with descriptive statistics.

```{r describe}
# DESCRIBE DATA
# options(digits=2)
itm_stats <- describe(all)
knitr::kable(itm_stats, caption = "Items Statistics", booktabs = TRUE, digits=2)

# A POSSIBLE BAR PLOT OF ITEM STATISTICS
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=TRUE, col='red')
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=FALSE, col='red', ylim = c(0,8000),names.arg=labs)

```

**Classical Item Analysis**.

```{r item-analysis}
# CLASSICAL ITEM ANALYSIS
# CTT::itemAnalysis(items, itemReport=TRUE, NA.Delete=TRUE, rBisML=FALSE, hardFlag, easyFlag, pBisFlag, bisFlag, flagStyle = c("X",""))
```


```{r}
# MORE ITEM STATISTICS
library(pastecs)
item_stats <- pastecs::stat.desc(all) # use the stat.desc() to render data frame of item stats
item_stats <- t(item_stats) # transpose the data frame
item_stats <- data.frame(item_stats)
item_stats[order(item_stats$mean), ] # order by mean
item_stats[order(item_stats$nbr.val), ] # order by N cases
```

**Distractor Analysis**.

```{r distractor}
# DISTRACTOR ANALYSIS
# CTT::distractorAnalysis
```

**Outlier Analysis**.

```{r outliers}
# OUTLIERS
# psych::outlier(items)
```

**Skew and Kurtosis**.

