# Item Analysis
In this chapter, we examine the technical quality of the items that Dan piloted for the SHAS. Here we address questions like the following:

- Which items, if any, are too "easy"? What is the impact of this?
- Which items, if any, are too "difficult"? What is the impact of this?
- Which items, if any, correlate weakly with others (and therefore might need to be refined or excluded from scaling?
- Which items, if any, have too many response categories which might be profitably collapsed?

We begin by defining item analysis and discussing its importance. Then we document our item analyses from an item response theory (IRT) framework. We conclude by summarizing what we learn from these analyses to help us refine these 66 items and possibly develop better ones in the future.

```{r, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(comment = "", warning = FALSE, echo = FALSE,  message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)
knitr::opts_knit$set(verbose = FALSE)
```

```{r}
# LOAD PACKAGES
library(lattice)
library(survival)
library(Formula)
library(ggplot2)
library(Hmisc) # for item statistics
library(psych) # for psychometric analysis
library(psychTools) # for psychometric analysis
library(CTT) # for classical test theory stats
library(pastecs) # for item statistics
library(ltm) # for point-biserial correlations
library(corrplot) # for correlation plot
library(knitr) # how RMarkdown compiles all the code into documents
```

```{r}
# LOAD DATA
items       <- read.csv("data/items.csv")
items_l     <- read.csv("data/items_l.csv") # listwise
```

## Item analyses in item response theory
Item response theory (IRT) is an established framework for empirically evaluating what items measure [@EmbretsonReise2000; @WrightMasters1982]. IRT methods complement methods derived from classical test theory in several ways.

First, IRT places both examinees and items on the same interval scale. In classical test theory, examinees differ in their scores and items differ in their means or percent of examinees selecting a response; examinees and items are on different scales. By contrast, IRT methods allow measurement of persons and items on the **_same scale_** with _equal interval properties_ of the _logit_ scale** and resulting linear measures (Wright & Stone, 1999). In IRT we can estimate _item parameters independently of the characteristics of the calibrating sample_, and we can estimate _person parameters apart from the difficulties of the items taken_ (Masters, 1982; Rasch, 1966).

Second, parameters for examinees and items are independent of each other.

Third, IRT gives items diagnostic value. A core assumption of CTT is that items are _redundant_; they contribute meaningful variation to a total score but they do not occupy a distinct place on the score continuum. IRT is different. IRT does not assume that items are redundant; instead it investigates whether items make distinct contributions to the measurement of the latent trait. Whereas classical test theory measurement is principally concerned with the ratio of noise to signal in the variation contributed by an item, IRT models compute the probability of a certain response to an item given the level of the latent construct the individual possesses (i.e., spiritual abuse) and the relevant item’s difficulty of endorsement. IRT is based on the logistic model and operates on the _logit_ scale.

In what follows, we conduct item analyses through a comparison of several different IRT models: the Rating Scale Model (RSM), the Partial Credit Model (PCM), the Generalized Partial Credit Model (GPCM). I credit Okan Bulut [-@Bulut2018] for his tutorial illustrating how to carry out IRT analyses in R using the `mirt` package [@mirt]. 

### Partial Credit Model (PCM)
The second model we examine is the Partial Credit Model (PCM) [@Masters1982]. This is a polytomous item response model belonging to the Rasch family. This means that it extends the Rasch model, which is based on dichotomous item responses (1=correct, 0=incorrect), by considering the probability of choosing from among several response options such as those typically presented in Likert scales of agreement. The number of response categories in each item may vary [@EmbretsonReise2000]. It follows that the PCM contains m (m + 1 being the number of response categories) threshold parameters, and each threshold parameter marks a category intersection [@Masters1982]. Threshold sometimes refers to “step difficulty” or “step parameter” and illustrates the point on the latent trait continuum (e.g., spiritual abuse) where a response in category k becomes more likely than a response in category k – 1 (de Ayala, 2009). The PCM provides scores for items, persons, and step parameters on a logit scale. We begin by loading the `mirt` library and estimating the model on the data from the 66 SHAS items. This model constrains the slopes to 1 and freely estimates the variance parameters.

```{r pcm-run}
# ESTIMATE PCM MODEL
library(stats4)
library(lattice)
library(mirt)
items <- read.csv("data/items.csv")
source("code/ApplyItemLabels.R")
model.pcm <- 'SPIRITUAL ABUSE PARTIAL CREDIT MODEL = 1-66'
results.pcm <- mirt(data=items, model=model.pcm, itemtype="Rasch", verbose=FALSE) # estimate model
```

```{r}
# GET ITEM PARAMETERS
coef.pcm <- coef(results.pcm, IRTpars=TRUE, simplify=TRUE) # save coefficients
items.pcm <- as.data.frame(coef.pcm$items) # item parameters as data frame
items.pcm$deltab1b2 <- items.pcm$b2 - items.pcm$b1 # look for reversals b1 to b2
items.pcm$deltab2b3 <- items.pcm$b3 - items.pcm$b2 # look for reversals b2 to b3
items.pcm$deltab3b4 <- items.pcm$b4 - items.pcm$b3 # look for reversals b3 to b4
rv <- subset(items.pcm, deltab1b2 < 0 | deltab2b3 < 0 | deltab3b4 <0) # subset items with reversals
rv2 <- subset(rv, select = -c(deltab1b2:deltab3b4))
```

```{r pcm-items, fig.cap='Partial Credit Model - Item Parameters - Items with Category Threshold Reversals'}
knitr::kable(rv2, digits = 2, caption = "Partial Credit Model Item Parameters - Items with Category Threshold Reversals")
```

**Item parameters**. Table \@ref(tab:pcm-items) reports the item parameters estimated by the PCM. The first column shows the discrimination parameter, which is equal to “1” for all items. This is because the Partial Credit Model, similar to the Rasch model, constrains the discrimination parameter to be “1”. The subsequent columns (b1 to b4) report the estimated thresholds or step parameters. Because the SHAS items all have five response categories, the PCM estimates four threshold parameters (b1 to b4) for each item. Ordered threshold parameters are not a requirement of the partial credit or generalized partial credit models [@EmbretsonReise2000]. The important question, therefore, is which items have ordered thresholds and which have "reversals." A reversal between two category thresholds suggests that one might not be contributing reliable information and might be collapsed.

As shown in Table \@ref(tab:pcm-items), most of the SHAS items (43 out of 66) had reversals. On each of these items, examinees who had experienced more overall lifetime spiritual abuse were more likely to report "Never" than "Once or twice".

**Item plots**. Item plots help show this information visually. The following figures are plots of the _option characteristic curves_ (also called category characteristic curves) of the first three items (Q7, Q10, Q11) with category reversals.

```{r pcm-a}
# PCM ITEM PLOTS
library(directlabels)
a <- plot(results.pcm, type = 'trace', which.items = 1,
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4))
direct.label(a, 'top.points')
```

```{r pcm-b}
# PCM ITEM PLOTS
b <- plot(results.pcm, type = 'trace', which.items = 2,
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4))
direct.label(b, 'top.points')
```

```{r pcm-c}
# PCM ITEM PLOTS
c <- plot(results.pcm, type = 'trace', which.items = 3,
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4))
direct.label(c, 'top.points')
```

In these plots, the x axis is the scale of the latent trait (lifetime exposure to spiritual abuse) expressed in logits ranging from -6 to +6. The y axis is the probability of selecting a given response (such as "Always") at each level of latent trait (overall lifetime exposure to spiritual abuse). The blue line is the probability of selecting "Never" for examinees ranging from low to high lifetime spiritual abuse. The reversal is that examinees just below zero are more likely to select "Never" than "Once or twice." In addition, examinees with above-average lifetime exposure to spiritual abuse were more likely to select "Always" than "Often". The item parameter statistics shown Table \@ref(tab:pcm-items) indicate that this issue applies to 43 items. These items might contribute more reliable information if these response categories were collapsed.

**Fit statistics**. Fit statistics gauge the extent to which the observed pattern of item responses fit the IRT model (Rasch, partial credit, etc.). They do this by means of chi-square statistics which examine the cumulative difference the observed pattern of item responses and the pattern of item respones that the model would expect. Two fit statistics commonly used in IRT modesl are the infit mean square (MNSQ) and the outfit mean square (MNSQ) [@BondFox2015]. The infit statistic places greater emphasis on unexpected responses that are close to the examinees and item location. The outfit is sensitive to unexpected responses that are far from the location. The expected value of infit or outfit for each item is 1.0, with a range of acceptable values ranging from 0.5 to 1.5. Values outside these boundaries indicate a lack of fit between items and the model.

```{r pcm-item-fits}
# ITEM FIT
ifit <- mirt::itemfit(results.pcm, na.rm=TRUE, 'infit')
i <- as.data.frame(itemfit(results.pcm, na.rm = TRUE, 'infit'))
misfits <- subset(i, outfit > 1.5 | outfit < 0.5 | infit > 1.5 | infit < 0.5)
```

```{r pcm-item-fit, fig.cap='Partial Credit Model (PCM) - Misfitting Items'}
knitr::kable(misfits, digits = 2, caption = "Partial Credit Model (PCM) - Misfitting Items", row.names = FALSE)
```

This information suggests a decent fit of the PCM to the SHAS item response data. All but six of the 66 items had infit and outfit statistics within the acceptable range from 0.5 to 1.5. Statistics for these six misfitting items are presented here in Table \@ref(tab:pcm-item-fit). The large outfit values mean the observed patterns of responses to these items differed substantially from the patterns of responses expected by the PCM. These items capture the experiences of women, people of color, and LGBTQ persons.

**Model fit**. 

```{r pcm-fit, fig.cap='PCM Model Fit'}
# MODEL FIT
M2(results.pcm, type = "C2", na.rm=TRUE, calcNULL = FALSE)
```

The obtained RMSEA value = .088 (95% CI[.088, .089]) and SRMSR value = .118 suggest that data do not fit the model reasonably very well using suggested cutoff values of RMSEA <= .06 and SRMSR <= .08 as suggested guidelines for assessing fit. The CFI = .961 exceeded the recommended .95 threshold.

