# Item Analysis
In this chapter, we examine the technical quality of the items that Dan piloted for the SHAS. Here we address questions like the following:

- Which items, if any, are too "easy"? What is the impact of this?
- Which items, if any, are too "difficult"? What is the impact of this?
- Which items, if any, correlate weakly with others (and therefore might need to be refined or excluded from scaling?
- Which items, if any, have too many response categories which might be profitably collapsed?

We begin by defining item analysis and discussing its importance. Then we document our item analyses from an item response theory (IRT) framework. We conclude by summarizing what we learn from these analyses to help us refine these 66 items and possibly develop better ones in the future.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

# LOAD PACKAGES
library(Hmisc) # for item statistics
library(psych) # for psychometric analysis
library(psychTools) # for psychometric analysis
library(CTT) # for classical test theory stats
library(pastecs) # for item statistics
library(ltm) # for point-biserial correlations
library(corrplot) # for correlation plot
library(knitr) # how RMarkdown compiles all the code into documents

# LOAD DATA
all         <- read.csv("data/all.csv")
items       <- read.csv("data/items.csv")
ext_items   <- read.csv("data/ext_items.csv")
int_items   <- read.csv("data/int_items.csv")
all_l       <- read.csv("data/all_l.csv") # listwise
items_l     <- read.csv("data/items_l.csv") # listwise
ext_items_l <- read.csv("data/ext_items_l.csv") # listwise
int_items_l <- read.csv("data/int_items_l.csv") # listwise
```

## Item analyses in item response theory
Item response theory (IRT) is an established framework for empirically evaluating what items measure [@EmbretsonReise2000; @WrightMasters1982]. IRT methods complement methods derived from classical test theory in several ways.

1. **Examinees and items on the same equal-interval scale**. In classical test theory, examinees differ in their scores and items differ in their means or percent of examinees selecting a response; examinees and items are on different scales. By contrast, IRT methods allow measurement of persons and items on the **_same scale_** with _equal interval properties_ of the _logit_ scale** and resulting linear measures (Wright & Stone, 1999). In IRT we can estimate _item parameters independently of the characteristics of the calibrating sample_, and we can estimate _person parameters apart from the difficulties of the items taken_ (Masters, 1982; Rasch, 1966).

2. **Item separation has diagnostic value**. A core assumption of CTT is that items are _redundant_; they contribute meaningful variation to a total score but they do not occupy a distinct place on the score continuum. IRT is different. IRT does not assume that items are redundant; instead it investigates whether items make distinct contributions to the measurement of the latent trait. 

Whereas classical test theory measurement is principally concerned with the ratio of noise to signal in the variation contributed by an item, IRT models compute the probability of a certain response to an item given the level of the latent construct the individual possesses (i.e., spiritual abuse) and the relevant item’s difficulty of endorsement. IRT is based on the logistic model and operates on the _logit_ scale.

In what follows, we conduct item analyses through a comparison of several different IRT models: the Partial Credit Model (PCM), the Generalized Partial Credit Model (GPCM), and the Rating Scale Model (RSM). I credit Okan Bulut [-@Bulut2020] for his excellent tutorial illustrating how to carry out IRT analyses in R using the `mirt` package (cite mirt). 

### IRT Model Assumptions
IRT makes two assumptions of the data. The first is unidimensionality of the latent trait. Is the trait one dimension? This is important for the SHAS because the questionnaire explicitly distinguished between items measuring external events and internal states. To check this assumption we first confirm the factor structure using confirmatory factor analysis (CFA) and Rasch Principal Component Analysis of Residuals (PCAR; Smith, 2002). 

The second assumption is local independence [@EmbretsonResise2000]. This means that any set of items should not share any meaningful correlation once the latent variable has been accounted for (Edwards, Houts, & Cai, 2018). 
To check the assumption of local independence, we used the Q3 statistic (Yen, 1984). The Q3 statistic index criteria are that the raw residual correlation between pairs of items should never exceed 0.10 (Marais & Andrich, 2008).

```{r q3, eval=FALSE, include = FALSE}

```

Using these criteria, we found XX items having local dependence. The items that had the highest raw residual correlations had negative signs and no positive residual correlation. In other words, the assumption of local independence in this study was met.

### Partial Credit Model (PCM)
The first model we examine is the Partial Credit Model (PCM) [@Masters1982]. This is a polytomous item response model belonging to the Rasch family. This means that it extends the Rasch model, which is based on dichotomous item responses (1=correct response, 0=incorrect response), by considering the probabliity of choosing from among several response options such as those typically presented in Likert scales of agreement or frequency. The PCM assumes that items use ordered response categories as they exist in questionnaires, such as unidimensional rating scales. The number of response categories in each item may vary [@EmbretsonReise2000]. It follows that the PCM contains m (m + 1 being the number of response categories) threshold parameters. Each threshold parameter marks a category intersection [@Masters1982]. Threshold sometimes refers to “step difficulty” or “step parameter” and illustrates the point on the latent trait continuum (e.g., spiritual abuse) where a response in category k becomes more likely than a response in category k – 1 (de Ayala, 2009). The PCM provides scores for items, persons, and step parameters on a logit scale.

**Model Estimation**. We begin by loading the `mirt` library and estimating the model on the data from the 66 SHAS items. This model constrains the slopes to 1 and freely estimates the variance parameters.

```{r mirtmod, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
# ESTIMATE MIRT MODEL
library(mirt)
items <- read.csv("data/items.csv")
model.pcm <- 'SPIRITUAL ABUSE = 1-66'
results.pcm <- mirt(data=items, model=model.pcm, itemtype="Rasch", verbose=FALSE)
```

```{r mirt-items}
coef.pcm <- coef(results.pcm, IRTpars=TRUE, simplify=TRUE)
items.pcm <- as.data.frame(coef.pcm$items)
# print(results.pcm)
knitr::kable(round(items.pcm,2), caption = "Partial Credit Model (PCM) - Item Parameters")
```
**Item parameters**. Table \@ref(tab:mirt-items) reports the item parameters estimated by the PCM. The first column shows the discrimination parameter, which is equal to “1” for all items because the Partial Credit Model, similar to the Rasch model, constrains the discrimination parameter to be “1”. In the following columns (b1 to b4) are the estimated thresholds or step parameters. Because the SHAS items all have five response categories, the Partial Credit Model estimates four threshold parameters (b1 to b4) for each item.  The question here is the spread of the parameters across the range of the trait.  In the PCM, "the property of ordered threshold parameters is not a requirement in the partial credit or generalized partial credit models" [@EmbretsonReise2000].

Consider the first item: Q7. The thresholds are as follows: b1=0.23, b2=0.08, b3=0.74, b4=1.11. The threshold for b1 is higher than b2. This means examinees who experienced more spiritual abuse overall were more likely to choose "Never" than "Once or twice" on this item.

**Item fit**. Two fit statistics assess the fit of an item to the PCM: the infit mean square (MNSQ) and the outfit mean square (MNSQ). The infit statistic places greater emphasis on unexpected responses that are close to the examinees and item location. The outfit is sensitive to unexpected responses that are far from the location [@BondFox2015]. The infit and outfit values identify potential unexpected response patterns. The expected value of infit or outfit for each item is 1.0, with a range of acceptable values ranging from 0.5 to 1.5. Values outside this boundary indicate a lack of fit between items and models.

```{r pcm-item-fit}
itemfit(results.pcm, na.rm=TRUE, 'infit')
```

As seen in the \@ref(tab:pcm-item-fit), XX items had infit and outfit statistics within an acceptable value (0.5 - 1.5). XX misfitting items were found and this means that XX items fit within the Rasch PCM. ^[and this is important because...]

**Item plots**. We can use also examine these item statistics visually. The following curves are called category response curves because they represent "the probability of an examinee (or survey respondent) responding in a particular category conditional on trait level" [@EmbretsonReise2000].  In general, "the higher the slope parameters, the steeper the operating characteristic curves and the more narrow and peaked the category response curves, indicating that the response categories differentiate among trait levels fairly well" [@EmbretsonReise2000].  Polytomous IRT models such as the Partial Credit Model have option characteristic curves (OCCs) which can be considered as an extension of item characteristic curves (ICCs) for polytomous items. Because the items have more than two response categories, OCCs have multiple curves in the same plot. Each curve represents the probability of selecting a particular response option as a function of the latent trait.

```{r pcm-occs}
# PCM ITEM PLOTS
# itemplot(results.pcm, 1, type='infotrace')
plot(results.pcm, type = 'trace', which.items = c(1,2,3,4,5,6,7,8,9,10,11,12),
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4)) # option curves
```

In Figure \@ref(fig:pcm-occs), each item has four OCCs representing the four response options. The response categories are labeled as “P1” to “P4”. For all four items, the OCCs follow the same order as the response categories. The OCC for the first response option (P1) is on the very left side of the plot, whereas the last response option (P4) is located on the right side of the plot.  There are two questions to ask of OCCs: (1) Are they in the correct order? and (2) Does each response category meaningfully represent the responses of respondents at a given trait level?  In this case, the OCCs are in the correct order, but the middle categories do not seem to capture distinct respondents.

**Model fit**. One way to evaluate the fit of the Partial Credit Model is to evaluate how well it fits the data from the item responses.  The following table presents this information.

```{r pcm-model-fit}
# PCM MODEL FIT
itemplot(results.pcm, 1, type = 'score')
empirical_plot(items, c(1,2,3,4), main = 'Empirical Plots') # multiple items
empirical_plot(items, 1, main = 'Empirical Plots') # one item
pcmfit <- as.data.frame(itemfit(results.pcm, na.rm=TRUE))
print(pcmfit, caption = 'Partial Credit Model (PCM) - Item Fit Statistics')
knitr::kable(pcmfit, caption = 'Partial Credit Model (PCM) - Item Fit Statistics')
itemfit(results.pcm, empirical.plot=1)
itemfit(results.pcm, 'infit', na.rm=TRUE)
```

Table X "contains estimation results and the fit index model from each item of the SHAS." The estimates of item difficulty were between -X.XX TO X.XX (ie., -1.36 to 1.51) on the logit scale. Based on the item discrimination index (PTBIS; point-biserial correlation), which is analogous to classical test theory (item to-total-correlation), X SHAS items have high discrimination values in a positive direction, with no values below 0.30 or negative values. This indicates that XX items function well to distinguish persons with high versus low levels of spiritual abuse. For each item, we examined the step parameter of category endorsements, and the statistics of each response for the items. X items, except item X, followed the step ordering requirement. We found Item X (Step 1 = -0.XX, Step 2 = -0.XX, Step 3 = 0.XX) experienced threshold disordering as the thresholds were not ordered from lowest to highest. However, none of the infit and outfit statistics for response categories were greater than 2. This finding indicated that collapsing categories was not necessary because the disordered threshold still fit and did not violate the Rasch model (Linacre, 2010; 2018). Based on this information, we concluded that, in general, the SHAS items (fit)/(did not fit) the Rasch PCM model.

### Generalized Partial Credit Model (GPCM)
In the PCM, "the property of ordered threshold parameters is not a requirement in the partial credit or generalized partial credit models" [@EmbretsonReise2000].

```{r gpcm-mod}
# GENERALIZED PCM ESTIMATION
model.gpcm <- 'SPIRITUAL ABUSE GPCM = 1-66'
results.gpcm <- mirt(data=items, model=model.gpcm, itemtype="gpcm", verbose=FALSE)
coef.gpcm <- coef(results.gpcm, IRTpars=TRUE, simplify=TRUE)
items.gpcm <- as.data.frame(coef.gpcm$items)
print(results.gpcm)
#print(items.gpcm)
knitr::kable(round(items.gpcm,2), caption = "Generalized Partial Credit Model (GPCM) - Item Parameters")
```

```{r gpcm-occs}
# OPTION CHARACTERISTIC CURVES
plot(results.gpcm, type = 'trace', which.items = c(1,2,3,4),
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4)) # option curves
```

**Item parameters**.

**Item plots**.

**Model fit**.

### Rating Scale Model (RSM)

```{r rsm-est}
model.rsm <- 'SPIRITUAL ABUSE RSM = 1-66'
results.rsm <- mirt(data=items, model=model.rsm, itemtype="rsm", verbose=FALSE)
coef.rsm <- coef(results.rsm, simplify=TRUE)
items.rsm <- as.data.frame(coef.rsm$items)
print(summary(results.rsm))
#print(items.rsm)
knitr::kable(round(items.rsm,2), caption = "Rating Scale Model (RSM) - Item Parameters")
```

```{r rsm-plots}
plot(results.rsm, type = 'trace', which.items = c(1,2,3,4),
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4)) # option curves
plot(results.rsm, type = 'itemscore', theta_lim = c(-4,4), lwd=2, facet_items = FALSE) # item scoring traceline

#plot(results.rsm, type = 'infotrace', which.items = c(1,2,3,4),
#     main = "", par.settings = simpleTheme(lwd=2)) # item information functions (IIF)
#plot(results.rsm, type = 'info', theta_lim = c(-4,4), lwd=2) # test information curve
#plot(results.rsm, type = 'SE', theta_lim = c(-4,4), lwd=2) # test standard errors
```


### Graded Response Model (GRM)

```{r}
model.grm <- 'SPIRITUAL ABUSE GRM = 1-66'
results.grm <- mirt(data=items, model=model.grm, itemtype="graded", verbose=FALSE)
coef.grm <- coef(results.grm, simplify=TRUE)
items.grm <- as.data.frame(coef.grm$items)
print(summary(results.grm))
#print(items.grm)
knitr::kable(round(items.grm,2), caption = "Graded Response Model (GRM) - Item Parameters")


plot(results.grm, type = 'trace', which.items = c(1,2,3,4),
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4)) # option curves
plot(results.grm, type = 'itemscore', theta_lim = c(-4,4), lwd=2, facet_items = FALSE) # item scoring traceline

#plot(results.rsm, type = 'infotrace', which.items = c(1,2,3,4),
#     main = "", par.settings = simpleTheme(lwd=2)) # item information functions (IIF)
#plot(results.rsm, type = 'info', theta_lim = c(-4,4), lwd=2) # test information curve
#plot(results.rsm, type = 'SE', theta_lim = c(-4,4), lwd=2) # test standard errors
```

In the Graded Response Model, between category threshold parameters are ordered.

## Summary
