# Item Analysis
This chapter documents our item analysis. This includes investigation of the following statistics from classical test theory:
- Item difficulty
- Item discrimination
- Item-total correlations

We also examine item statistics from item response theory (IRT) including:
- Item location parameters
- Item characteristic curves (ICCs)

The goal of this information with which to improve items. One outcome of this analysis is insight into whether items have too many response categories.

## Preliminary work

**Load packages**. This code chunk outlines the R packages we use and the data. There are several helpful packages for classical test theory (CTT).

```{r, message=FALSE, warning=FALSE}
# LOAD PACKAGES
library(Hmisc) # for item statistics
library(psych) # for psychometric analysis
library(psychTools) # for psychometric analysis
library(CTT) # for classical test theory stats
library(pastecs) # for item statistics
library(ltm) # for point-biserial correlations
library(corrplot) # for correlation plot
library(knitr) # how RMarkdown compiles all the code into documents
```

**Load data**. This code chunk outlines the R packages we use and the data. There are several helpful packages for classical test theory (CTT).

```{r}
# LOAD DATA

# Datasets straight off the Excel file
all         <- read.csv("data/all.csv")
items       <- read.csv("data/items.csv")
ext_items   <- read.csv("data/ext_items.csv")
int_items   <- read.csv("data/int_items.csv")

# Datasets with listwise deletion
all_l       <- read.csv("data/all_l.csv")
# items_l     <- read.csv("data/items_l.csv")
# ext_items_l <- read.csv("data/ext_items_l.csv")
# int_items_l <- read.csv("data/int_items_l.csv")
```

##Classical Item Analysis


```{r describe, fig.cap='Item Statistics'}
# DESCRIBE DATA
# options(digits=2)
# describe(items)

# this doesn't work
# knitr::kable(itm_stats, caption = "Items Statistics", booktabs = TRUE, digits=2)

# A POSSIBLE BAR PLOT OF ITEM STATISTICS
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=TRUE, col='red')
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=FALSE, col='red', ylim = c(0,8000),names.arg=labs)

```


```{r item-analysis}
# CLASSICAL ITEM ANALYSIS

# CTT::itemAnalysis(items, itemReport=TRUE, NA.Delete=TRUE, rBisML=FALSE, hardFlag, easyFlag, pBisFlag, bisFlag, flagStyle = c("X",""))

CTT::itemAnalysis(items, itemReport=TRUE, NA.Delete=TRUE, rBisML=FALSE, flagStyle = c("X",""))
```

```{r}
# MORE ITEM STATISTICS
library(pastecs)
item_stats <- pastecs::stat.desc(items) # use the stat.desc() to render data frame of item stats
item_stats <- t(item_stats) # transpose the data frame
item_stats <- data.frame(item_stats)
item_stats[order(item_stats$mean), ] # order by mean
item_stats[order(item_stats$nbr.val), ] # order by N cases
```

**Distractor Analysis**.

```{r distractor}
# DISTRACTOR ANALYSIS
# CTT::distractorAnalysis
```

**Outlier Analysis**.

```{r outliers}
# OUTLIERS
# psych::outlier(items)
```


## Item response theory

This ltm package "provides a flexible framework for Item Response Theory analyses for dichotomous
and polytomous data under a Marginal Maximum Likelihood approach. The fitting algorithms
provide valid inferences under Missing At Random missing data mechanisms" [@ltm].

```{r ext-pbis, fig.cap='Item Total Correlations for External Events Items'}
# DESCRIPTIVES
ltm::descript(items)
```






**Skew and Kurtosis**.

