# Item Analysis
In this chapter, we examine the technical quality of the 66 items that Dan piloted for the SHAS. We begin by defining item analysis and discussing its importance. Then document our item analyses from both classical test theory (CTT) and item response theory (IRT) frameworks. We conclude by summarizing what we learn from these analyses to help us refine these items and develop better ones in the future.

## About Item Analysis


```{r, message=FALSE, warning=FALSE}

# LOAD PACKAGES
library(Hmisc) # for item statistics
library(psych) # for psychometric analysis
library(psychTools) # for psychometric analysis
library(CTT) # for classical test theory stats
library(pastecs) # for item statistics
library(ltm) # for point-biserial correlations
library(corrplot) # for correlation plot
library(knitr) # how RMarkdown compiles all the code into documents

# LOAD DATA
all         <- read.csv("data/all.csv")
items       <- read.csv("data/items.csv")
ext_items   <- read.csv("data/ext_items.csv")
int_items   <- read.csv("data/int_items.csv")
# Datasets with listwise deletion
all_l       <- read.csv("data/all_l.csv")
# items_l     <- read.csv("data/items_l.csv")
# ext_items_l <- read.csv("data/ext_items_l.csv")
# int_items_l <- read.csv("data/int_items_l.csv")


```

## Classical Item Analysis

```{r describe, fig.cap='Item Statistics'}
# DESCRIBE DATA
# options(digits=2)
# describe(items)

# this doesn't work
# knitr::kable(itm_stats, caption = "Items Statistics", booktabs = TRUE, digits=2)

# A POSSIBLE BAR PLOT OF ITEM STATISTICS
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=TRUE, col='red')
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=FALSE, col='red', ylim = c(0,8000),names.arg=labs)

```


```{r item-analysis}
# CLASSICAL ITEM ANALYSIS

# CTT::itemAnalysis(items, itemReport=TRUE, NA.Delete=TRUE, rBisML=FALSE, hardFlag, easyFlag, pBisFlag, bisFlag, flagStyle = c("X",""))

CTT::itemAnalysis(items, itemReport=TRUE, NA.Delete=TRUE, rBisML=FALSE, flagStyle = c("X",""))
```

```{r}
# MORE ITEM STATISTICS
library(pastecs)
item_stats <- pastecs::stat.desc(items) # use the stat.desc() to render data frame of item stats
item_stats <- t(item_stats) # transpose the data frame
item_stats <- data.frame(item_stats)
item_stats[order(item_stats$mean), ] # order by mean
item_stats[order(item_stats$nbr.val), ] # order by N cases
```

**Distractor Analysis**.

```{r distractor}
# DISTRACTOR ANALYSIS
# CTT::distractorAnalysis
```

**Outlier Analysis**.

```{r outliers}
# OUTLIERS
# psych::outlier(items)
```

**Skew and Kurtosis**.

## Item analyses in item response theory (IRT) framework

One outcome of this analysis is insight into whether items have too many response categories.

This ltm package "provides a flexible framework for Item Response Theory analyses for dichotomous and polytomous data under a Marginal Maximum Likelihood approach. The fitting algorithms provide valid inferences under Missing At Random missing data mechanisms" [@ltm].

```{r ext-pbis, fig.cap='Item Total Correlations for External Events Items'}
# DESCRIPTIVES
ltm::descript(items)
```

### Item thresholds

### Item characteristic curves


### Differential item functioning (DIF)

## Summary
