# Item Analysis
In this chapter, we examine the technical quality of the items that Dan piloted for the SHAS. Here we address questions like the following:

- Which items, if any, are too "easy"? What is the impact of this?
- Which items, if any, are too "difficult"? What is the impact of this?
- Which items, if any, correlate weakly with others (and therefore might need to be refined or excluded from scaling?
- Which items, if any, have too many response categories which might be profitably collapsed?

We begin by defining item analysis and discussing its importance. Then we document our item analyses from an item response theory (IRT) framework. We conclude by summarizing what we learn from these analyses to help us refine these 66 items and possibly develop better ones in the future.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

# LOAD PACKAGES
library(Hmisc) # for item statistics
library(psych) # for psychometric analysis
library(psychTools) # for psychometric analysis
library(CTT) # for classical test theory stats
library(pastecs) # for item statistics
library(ltm) # for point-biserial correlations
library(corrplot) # for correlation plot
library(knitr) # how RMarkdown compiles all the code into documents

# LOAD DATA
all         <- read.csv("data/all.csv")
items       <- read.csv("data/items.csv")
ext_items   <- read.csv("data/ext_items.csv")
int_items   <- read.csv("data/int_items.csv")
all_l       <- read.csv("data/all_l.csv") # listwise
items_l     <- read.csv("data/items_l.csv") # listwise
ext_items_l <- read.csv("data/ext_items_l.csv") # listwise
int_items_l <- read.csv("data/int_items_l.csv") # listwise
```

## Item analyses in item response theory
Item response theory (IRT) is an established framework for empirically evaluating what items measure [@EmbretsonReise2000; @WrightMasters1982]. IRT methods complement methods derived from classical test theory in several ways.

First, IRT places both examinees and items on the same interval scale. In classical test theory, examinees differ in their scores and items differ in their means or percent of examinees selecting a response; examinees and items are on different scales. By contrast, IRT methods allow measurement of persons and items on the **_same scale_** with _equal interval properties_ of the _logit_ scale** and resulting linear measures (Wright & Stone, 1999). In IRT we can estimate _item parameters independently of the characteristics of the calibrating sample_, and we can estimate _person parameters apart from the difficulties of the items taken_ (Masters, 1982; Rasch, 1966).

Second, IRT gives items diagnostic value. A core assumption of CTT is that items are _redundant_; they contribute meaningful variation to a total score but they do not occupy a distinct place on the score continuum. IRT is different. IRT does not assume that items are redundant; instead it investigates whether items make distinct contributions to the measurement of the latent trait. Whereas classical test theory measurement is principally concerned with the ratio of noise to signal in the variation contributed by an item, IRT models compute the probability of a certain response to an item given the level of the latent construct the individual possesses (i.e., spiritual abuse) and the relevant item’s difficulty of endorsement. IRT is based on the logistic model and operates on the _logit_ scale.

In what follows, we conduct item analyses through a comparison of several different IRT models: the Partial Credit Model (PCM), the Generalized Partial Credit Model (GPCM), and the Rating Scale Model (RSM). I credit Okan Bulut [-@Bulut2018] for his tutorial illustrating how to carry out IRT analyses in R using the `mirt` package [@mirt]. 

### IRT Model Assumptions
IRT makes two assumptions of the data. The first is unidimensionality of the latent trait. Is the trait one dimension? This is important for the SHAS because the questionnaire explicitly distinguished between items measuring external events and internal states. To check this assumption we first confirm the factor structure using confirmatory factor analysis (CFA) and Rasch Principal Component Analysis of Residuals (PCAR; Smith, 2002). The second assumption is local independence [@EmbretsonReise2000]. This means that any set of items should not share any meaningful correlation once the latent variable has been accounted for (Edwards, Houts, & Cai, 2018).

To check the assumption of local independence, we used the Q3 statistic (Yen, 1984). The Q3 statistic index criteria are that the raw residual correlation between pairs of items should never exceed 0.10 (Marais & Andrich, 2008).

```{r q3, eval=FALSE, include = FALSE}

```

Using these criteria, we found XX items in possession of local dependence. The items that had the highest raw residual correlations had negative signs and no positive residual correlation. In other words, the assumption of local independence in this study (was)/(was not) met.

### Partial Credit Model (PCM)
The first model we examine is the Partial Credit Model (PCM) [@Masters1982]. This is a polytomous item response model belonging to the Rasch family. This means that it extends the Rasch model, which is based on dichotomous item responses (1=correct, 0=incorrect), by considering the probability of choosing from among several response options such as those typically presented in Likert scales of agreement. The number of response categories in each item may vary [@EmbretsonReise2000]. It follows that the PCM contains m (m + 1 being the number of response categories) threshold parameters, and each threshold parameter marks a category intersection [@Masters1982]. Threshold sometimes refers to “step difficulty” or “step parameter” and illustrates the point on the latent trait continuum (e.g., spiritual abuse) where a response in category k becomes more likely than a response in category k – 1 (de Ayala, 2009). The PCM provides scores for items, persons, and step parameters on a logit scale. We begin by loading the `mirt` library and estimating the model on the data from the 66 SHAS items. This model constrains the slopes to 1 and freely estimates the variance parameters.

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# ESTIMATE PCM MODEL
library(mirt)
items <- read.csv("data/items.csv")
source("code/ApplyItemLabels.R")
model.pcm <- 'SPIRITUAL ABUSE PARTIAL CREDIT MODEL = 1-66'
results.pcm <- mirt(data=items, model=model.pcm, itemtype="Rasch", verbose=FALSE) # estimate model
coef.pcm <- coef(results.pcm, IRTpars=TRUE, simplify=TRUE) # save coefficients
items.pcm <- as.data.frame(coef.pcm$items) # item parameters as data frame
items.pcm$deltab1b2 <- items.pcm$b2 - items.pcm$b1 # look for reversals b1 to b2
items.pcm$deltab2b3 <- items.pcm$b3 - items.pcm$b2 # look for reversals b2 to b3
items.pcm$deltab3b4 <- items.pcm$b4 - items.pcm$b3 # look for reversals b3 to b4
rv <- subset(items.pcm, deltab1b2 < 0 | deltab2b3 < 0 | deltab3b4 <0) # subset items with reversals
```

```{r pcm-items, fig.cap='Partial Credit Model - Item Parameters - Items with Category Threshold Reversals'}
knitr::kable(rv, "simple", digits = 2, caption = "Partial Credit Model - Item Parameters - Items with Category Threshold Reversals", full = TRUE)
```

Table \@ref(tab:pcm-items) reports the item parameters estimated by the PCM. The first column shows the discrimination parameter, which is equal to “1” for all items. This is because the Partial Credit Model, similar to the Rasch model, constrains the discrimination parameter to be “1”. The subsequent columns (b1 to b4) report the estimated thresholds or step parameters. Because the SHAS items all have five response categories, the PCM estimates four threshold parameters (b1 to b4) for each item. Ordered threshold parameters are not a requirement of the partial credit or generalized partial credit models [@EmbretsonReise2000]. The important question, therefore, is which items have ordered thresholds and which have "reversals." A reversal suggests that a response category is not contributing reliable information and might be collapsed.

As shown in Table \@ref(tab:pcm-items), most of the SHAS items (43 out of 66) had reversals. On each of these items, examinees who had experienced more overall lifetime spiritual abuse were more likely to report "Never" than "Once or twice". To collapse these categories would help these items contribute less noise and more reliable information.

Item plots help illustrate these statistics visually. Figure \@ref(pcm-occs) is a plot of the option characeristic curves (also called category characteristic curves) of Q7.

```{r pcm-occs}
# PCM ITEM PLOTS
library(directlabels)
plt <- plot(results.pcm, type = 'trace', which.items = 3,
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4)) # option curves
direct.label(plt, 'top.points')
```
In this kind of plot, the x axis is the scale of the latent trait--in this case, lifetime exposure to spiritual abuse--expressed in logits ranging from -6 to +6. The y axis is the probability of selecting a given response (such as "Always") at each level of overall lifetime exposure to spiritual abuse. The blue line is the thus the probability of selecting "Never" for examinees ranging from low to high lifetime spiritual abuse. The reversal is that examinees just below zero are more likely to select "Never" than "Once or twice." In addition, examinees with above-average lifetime exposure to spiritual abuse were more likely to select "Always" than "Often". The item parameter statistics shown Table \@ref(tab:pcm-items) indicate that this issue applies to 43 items. These items would contribute less noise and more reliable information if these response categories were collapsed.

**Item fit**. Two fit statistics assess the fit of an item to the PCM: the infit mean square (MNSQ) and the outfit mean square (MNSQ). The infit statistic places greater emphasis on unexpected responses that are close to the examinees and item location. The outfit is sensitive to unexpected responses that are far from the location [@BondFox2015]. The infit and outfit values identify potential unexpected response patterns. The expected value of infit or outfit for each item is 1.0, with a range of acceptable values ranging from 0.5 to 1.5. Values outside this boundary indicate a lack of fit between items and models.

```{r echo=FALSE, eval=TRUE, include=FALSE, warning=FALSE, message=FALSE}
# ITEM FIT
ifit <- itemfit(results.pcm, na.rm=TRUE, 'infit')
# itemfit(results.pcm, na.rm = TRUE, 'infit')
i <- as.data.frame(itemfit(results.pcm, na.rm = TRUE, 'infit'))
misfits <- subset(i, outfit > 1.5 | outfit < 0.5 | infit > 1.5 | infit < 0.5)
```

```{r pcm-item-fit, fig.cap='Partial Credit Model (PCM) - Misfitting Items'}
knitr::kable(misfits, "simple", digits = 2, caption = "Partial Credit Model (PCM) - Misfitting Items")
```

All but six of the 66 SHAS items had infit and outfit statistics within the acceptable range from 0.5 to 1.5. Statistics for these six items are presented here in \@ref(tab:pcm-item-fit). The large outfit values mean the observed patterns of responses to these items differed substantially from the patterns of responses expected by the PCM. These items capture the experiences of women, people of color, and LGBTQ persons.

**Model fit**. One way to evaluate the fit of the Partial Credit Model is to evaluate how well it fits the data from the item responses.  The following table presents this information.

```{r pcm-model-fit}
# PCM MODEL FIT
itemplot(results.pcm, 1, type = 'score')
empirical_plot(items, c(1,2,3,4), main = 'Empirical Plots') # multiple items
empirical_plot(items, 1, main = 'Empirical Plots') # one item
pcmfit <- as.data.frame(itemfit(results.pcm, na.rm=TRUE))
knitr::kable(pcmfit, caption = 'Partial Credit Model (PCM) - Item Fit Statistics')
itemfit(results.pcm, empirical.plot=1)
itemfit(results.pcm, 'infit', na.rm=TRUE)
```

Table X "contains estimation results and the fit index model from each item of the SHAS." The estimates of item difficulty were between -X.XX TO X.XX (ie., -1.36 to 1.51) on the logit scale. Based on the item discrimination index (PTBIS; point-biserial correlation), which is analogous to classical test theory (item to-total-correlation), X SHAS items have high discrimination values in a positive direction, with no values below 0.30 or negative values. This indicates that XX items function well to distinguish persons with high versus low levels of spiritual abuse. For each item, we examined the step parameter of category endorsements and the statistics of each response for the items. X items, except item X, followed the step ordering requirement. 

We found Item X (Step 1 = -0.XX, Step 2 = -0.XX, Step 3 = 0.XX) experienced threshold disordering as the thresholds were not ordered from lowest to highest. However, none of the infit and outfit statistics for response categories were greater than 2. This finding indicated that collapsing categories was not necessary because the disordered threshold still fit and did not violate the Rasch model (Linacre, 2010; 2018). Based on this information, we concluded that, in general, the SHAS items (fit)/(did not fit) the Rasch PCM model.

```{r}
pcmscores <- fscores(results.pcm)
items <- cbind(items, pcmscores)
```

## Summary
