# Dimensionality
This chapter deals with the topic of dimensionality. It begins with a definition of dimensionality and explains how it is important. It outlines how to investigate it and suggests what counts as good evidence. Then it reports our study of the dimensionality of the SHAS.

## About Dimensionality
Dimensionality is the study of the dimensions (also known as "factors", "components", and "latent traits") that the items are designed and assembled to collectively measure. This study tends to take two forms. One is exploration of what dimensions a set of items apparently measure, regardless of what they intended to measure. The other is what evidence exist to confirm that a set of items actually measure the dimension they are intended to measure. 

Dimensionality is important for the reporting of scores. Scores, as summary measures of a set of items, are only meaningful to the extent that they enjoy evidence that they measure the dimension(s) they are designed to measure. Scores, in turn, are important to the extent that they support decisions that have consequences in the real world. Scores are needed to decide who gets to practice medicine, or fly an airplane. 

Dimensionality is an important part of the validation process for psychological measures [@EmbretsonReise2000].

The methodology commonly used to investigate dimensionality is factor analysis. To assess the dimensionality of these 66 piloted SHAS items we use exploratory factor analysis and confirmatory factor analysis (CFA) and PCAR.

## Exploratory Factor Analysis (EFA)

```{r}
# LOAD PACKAGES
library(lavaan)
library(psych)
library(psychTools)

# LOAD DATA
# all         <- read.csv("data/all.csv")
items       <- read.csv("data/items.csv")
ext_items   <- read.csv("data/ext_items.csv")
int_items   <- read.csv("data/int_items.csv")

```

```{r}
# TEST FOR NUMBER OF FACTORS IN DATA USING PARALLEL ANALYSIS AND VERY SIMPLE STRUCTURE (VSS)
fa.parallel(items)
vss(items)
```

```{r}
# PRINCIPAL COMPONENTS ANALYSIS
principal(items)
```

```{r}
# PRINCIPAL AXIS FACTORING
# library(psych)
# fit <- factor.pa(items, nfactors=3, rotation="varimax")
# fit # print results
```

```{r}
# PRINCIPAL COMPONENTS ANALYSIS
# entering raw data and extracting PCs
# from the correlation matrix
# fit <- princomp(items, cor=TRUE)
# summary(fit) # print variance accounted for
# loadings(fit) # pc loadings
# plot(fit,type="lines") # scree plot
# fit$scores # the principal components
# biplot(fit)
```

```{r}
# DETERMINE NUMBER OF FACTORS TO EXTRACT
# library(nFactors)
# ev <- eigen(cor(items)) # get eigenvalues
# ap <- parallel(subject=nrow(items),var=ncol(items),
#   rep=100,cent=.05)
# nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
# plotnScree(nS)
```

```{r}
# MAXIMUM LIKELIHOOD FACTOR ANALYSIS
# entering raw data and extracting 3 factors,
# with varimax rotation
# fit <- factanal(items, 3, rotation="varimax")
# print(fit, digits=2, cutoff=.3, sort=TRUE)
# # plot factor 1 by factor 2
# load <- fit$loadings[,1:2]
# plot(load,type="n") # set up plot
# text(load,labels=names(items),cex=.7) # add variable names

```

## Confirmatory Factor Analysis (CFA)
The purpose of confirmatory factor analysis is, as the name implies, to confirm that items measure the factor(s) or trait(s) for which the items were designed and intended to measure.

In our use of CFA we used several statistics and fit indices:

- Root Mean Square Error of Approximation (RMSEA)
- Comparative Fit Index (CFI)
- Tucker–Lewis Index (TLI), and
- Standardized Root-Mean-Square Residual (SRMR)

Based on published criteria (Hu & Bentler, 1999; Wang & Wang, 2019), we used the following standards for good fit:

- CFI > 0.95
- TLI > 0.95
- RMSEA < 0.05
- SRMR < 0.08

We use the MLM estimator [@Brown2006].

```{r}
# ltm::factor.scores(items) # Computation of factor scores for grm, ltm, rasch and tpm models
```

```{r}
# library(lavaan)
# MODEL WITH ONE COMMON FACTOR

# wb_1factor <- ' #start of model

# latent variable definitions (common factors)
  # wellbeing =~ NA*sense.peace + 
  #           NA*sense.wonder +
  #           NA*sense.gratitude +
  #           NA*sense.purpose
  
# latent variable variances
  # wellbeing ~~ 1*wellbeing

# latent variable covariances
# latent variable means

# manifest variable variances (uniquenesses)
  # sense.peace ~~ sense.peace
  # sense.wonder ~~ sense.wonder
  # sense.gratitude ~~ sense.gratitude
  # sense.purpose ~~ sense.purpose
            
# manifest variable covariances (uniquenesses)

#manifest variable means 
  # sense.peace ~ 1
  # sense.wonder ~ 1
  # sense.gratitude ~ 1
  # sense.purpose ~ 1
 # end of model

# CFA1 <- lavaan(wb_1factor, data=items, estimator = "MLM") # estimate model
# summary(CFA1, standardized=TRUE, fit.measures=TRUE)
```

The results of CFA, using MLR estimation, confirmed the second order factor model, in line with the original structure of the scale (see Olufadi, 2017), because the values of the indices were above the acceptable threshold [χ2 (167) = 284.032, p < 0.001; RMSEA = 0.030 (90% CI = 0.024-0.036), CFI = 0.983, TLI = 0.981, SRMR = 0.023], compared to the unidimensional model [χ2 (170) = 316.949, p < 0.001; RMSEA = 0.034 (90% CI = 0.028-0.039), CFI = 0.978, TLI = 0.976, SRMR = 0.023] and a 3-uncorrelated-factors model [χ2 (170) = 2427.628, p < 0.001; RMSEA = 0.132 (90% CI = 0.127-0.136), CFI = 0.669, TLI = 0.631, SRMR = 0.345]. Based on RMSEA, CFI, TLI, and SRMR, the results indicated that the model provided satisfactory representations of the underlying structure of the psychological well-being construct. All items loaded significantly (ranging from 0.535 to 0.817) in relation to each first order factor, at a p < 0.01 significance level.  These results of the CFA offer evidence that the four items measure one dimension of well-being.

## Summary
