# Confirmatory Factor Analysis

## Confirmatory Factor Analysis (CFA)
The purpose of confirmatory factor analysis is, as the name implies, to confirm that items measure the factor(s) or trait(s) for which the items were designed and intended to measure.

In our use of CFA we used several statistics and fit indices:

- Root Mean Square Error of Approximation (RMSEA)
- Comparative Fit Index (CFI)
- Tucker–Lewis Index (TLI), and
- Standardized Root-Mean-Square Residual (SRMR)

Based on published criteria (Hu & Bentler, 1999; Wang & Wang, 2019), we used the following standards for good fit:

- CFI > 0.95
- TLI > 0.95
- RMSEA < 0.05
- SRMR < 0.08

We use the MLM estimator [@Brown2006].

```{r}
# ltm::factor.scores(items) # Computation of factor scores for grm, ltm, rasch and tpm models
```

```{r}
# library(lavaan)
# MODEL WITH ONE COMMON FACTOR

# wb_1factor <- ' #start of model

# latent variable definitions (common factors)
  # wellbeing =~ NA*sense.peace + 
  #           NA*sense.wonder +
  #           NA*sense.gratitude +
  #           NA*sense.purpose
  
# latent variable variances
  # wellbeing ~~ 1*wellbeing

# latent variable covariances
# latent variable means

# manifest variable variances (uniquenesses)
  # sense.peace ~~ sense.peace
  # sense.wonder ~~ sense.wonder
  # sense.gratitude ~~ sense.gratitude
  # sense.purpose ~~ sense.purpose
            
# manifest variable covariances (uniquenesses)

#manifest variable means 
  # sense.peace ~ 1
  # sense.wonder ~ 1
  # sense.gratitude ~ 1
  # sense.purpose ~ 1
 # end of model

# CFA1 <- lavaan(wb_1factor, data=items, estimator = "MLM") # estimate model
# summary(CFA1, standardized=TRUE, fit.measures=TRUE)
```

The results of CFA, using MLR estimation, confirmed the second order factor model, in line with the original structure of the scale (see Olufadi, 2017), because the values of the indices were above the acceptable threshold [χ2 (167) = 284.032, p < 0.001; RMSEA = 0.030 (90% CI = 0.024-0.036), CFI = 0.983, TLI = 0.981, SRMR = 0.023], compared to the unidimensional model [χ2 (170) = 316.949, p < 0.001; RMSEA = 0.034 (90% CI = 0.028-0.039), CFI = 0.978, TLI = 0.976, SRMR = 0.023] and a 3-uncorrelated-factors model [χ2 (170) = 2427.628, p < 0.001; RMSEA = 0.132 (90% CI = 0.127-0.136), CFI = 0.669, TLI = 0.631, SRMR = 0.345]. Based on RMSEA, CFI, TLI, and SRMR, the results indicated that the model provided satisfactory representations of the underlying structure of the psychological well-being construct. All items loaded significantly (ranging from 0.535 to 0.817) in relation to each first order factor, at a p < 0.01 significance level.  These results of the CFA offer evidence that the four items measure one dimension of well-being.