# Data Preparation
This chapter documents our decisions surrounding how to prepare the data for analysis. Here we get into the weeds. We begin with loading the data from the original data file. From this we select a subset of just the items to which we apply labels and recode so that they begin from 0. We deal with missing data. The chapter concludes with a final set of datasets for use in the remaining chapters.

## Load data
The data are contained in an Excel file, simply named "data.xlsx", in the "data" folder of this project. The important thing is to keep all the data in the same folder within the larger "SHAS Validity" project folder.

```{r load-data, include=TRUE, echo=TRUE}
# LOAD ORIGINAL DATA
library(readxl) # for reading the source Excel file
shas <- read_excel("data/data.xlsx") # load Excel file
```

## Select items
Which items should we include for analysis? The first 13 columns in the original data file report data from several screening items that were placed at the beginning of the questionnaire. But because they are constant and not variable, they don't seem to have a place in scale development. Therefore it seems the important columns range from columns 14 to 97. The first set of items measure the **External Events** scale. These 52 items range from columns 14 to 35. The second set of items measure the **Internal States** scale, and these 14 items range from columns 66 to 79. The last set of items capture **demographic** variables, and these range from columns 80 to 97.

```{r select-items}
# SELECT ITEMS
all <- shas[ ,14:97] # select items, dems from columns 14-79
all <- subset(all, select = -c(Ext_Avg:Global_Avg)) # remove columns for now
items <- all[ ,1:66]
```

## Select item clusters
We select two subsets of the item data, one from just the External Events items and the other from the Internal States items.

```{r subsets}
ext_items <- items[ ,1:52]
int_items <- items[ ,52:66]
```

## Dichotomized data
Some IRT models (such as the Rasch model) and their estimation procedures in various R packages are based on dichotomous data (0=incorrect, 1=correct) rather than the polytomous data that comes from Likert scales. It is also possible that the extra categories create noise and improved SHAS items might use fewer response categories. Here we create several data sets of dichotomized SHAS items.

```{r dichot}
# DICHOTOMIZED DATA SETS
ditems <- items
ditems[ditems == 1] <- 0
ditems[ditems == 2] <- 1
ditems[ditems == 3] <- 1
ditems[ditems == 4] <- 1
ditems[ditems == 5] <- 1

dext_items <- ditems[ ,1:52]
dint_items <- ditems[ ,52:66]
```

## Missing data
Missing data can be an important problem. The problem has two parts. The first is how many data are missing. Obviously the more data are missing, the larger the barrier this problem poses to analysis. The second is the mechanism of missing data. Are the missing data randomly, or is there a pattern to the missing data - like a problematic item, or a distinct group of examinees who chose not to respond? Figure \@ref(fig:vim-plots) presents two plots from the VIM package for missing data [@R-VIM]. 

```{r vim-plots, fig.cap='Missing Data', message=FALSE, warning=FALSE}
# VISUALIZE AND QUANTIFY MISSING DATA
library(VIM)
a <- VIM::aggr(all, plot = FALSE)
plot(a, numbers = TRUE, prop = FALSE)
```

The first plot in Figure \@ref(fig:vim-plots) is a frequency distribution of the number of missing data entries for each item. This plot reveals very few missing data for the first set of External Events items, but nearly 200 missing entries for the second set of Internal States items and demographics. The raises two questions. The first is: Who were those examinees? We know that they responded to the first 52 items, then apparently quit the survey. Why? Was it just fatigue, or is there more to know about them? The second question is what to do with the information they actually did provide.

The second plot in Figure \@ref(fig:vim-plots) illustrates the different combinations of missing entries. The blue squares represent cases with complete responses, and the red indicates missing data. The rows represent combinations of complete and missing responses. The bottom row of entirely blue squares represents complete data -- examinees who responded to all of the items. The barplot on the right axis indicates that these were the most frequent combinations. The second most frequent combination was examinees who completed all of the External Events items and then quit the survey. Beyond that, there is no apparent order or pattern to the missing data.

```{r eval=FALSE, include=FALSE}
# VISUALIZE AND QUANTIFY MISSING DATA
# library(naniar)
# naniar::vis_miss(all)
```

**Mechanism of missing data**. @McKnightEtAl2007 prescribe several approaches for investigating the mechanisms of missing data. An appropriate first step is to test whether the data are Missing Completely At Random (MCAR) with the use of Little's test @Little1988. This is a chi-square test of the difference between the observed pattern of missing responses and the pattern of responses that would expected if missingness were completely random. We report the results of this test in Table \@ref(tab:mcar).

```{r mcar, message=FALSE, warning=FALSE}
library(naniar)
library(knitr)
knitr::kable(mcar_test(all), digits = 2, caption = 'Little\'s Test of Missing Completely at Random (MCAR)')
```

The test is fairly conclusive. The p-value for the chi-square is 1, which means observed pattern of missing data do not diverge significantly from what would be expected if the data were missing completely at random.

**Missing data methods**. There are several different ways to handle missing data. A common one is to _delete all cases missing any data_. This is called **listwise** deletion of data. Its primary advantages are simplicity, expedience, and analysis based on complete information. Its disadvantages are (1) disposal of the information offered by those examinees who responded to some items but not others, and (2) diminishment of the dataset, and with it, a reduction in statistical power. There are also methods of imputing values to missing responses. Here, for expedience, we use listwise deletion to create several datasets with complete data.

```{r listwise, include=TRUE, warning=FALSE, tidy=TRUE}
# LISTWISE DELETION OF MISSING DATA
all_l <- na.omit(all) #  All items + demographics
items_l <- na.omit(items) # All Likert items
ext_items_l <- na.omit(ext_items) # Likert items for External Events post listwise
int_items_l <- na.omit(int_items) # Likert items for Internal States post listwise
ditems_l <- na.omit(ditems) # All dichotomized Likert items post listwise
dext_items_l <- na.omit(dext_items) # Dichotomized External Events items post listwise
dint_items_l <- na.omit(dint_items) # Dichotomized Internal States items post listwise
```

## Final datasets

```{r save-data}
# SAVE DATA FILES
write.csv(all,"data/all.csv", row.names=FALSE)
write.csv(items,"data/items.csv", row.names=FALSE)
write.csv(ext_items,"data/ext_items.csv", row.names=FALSE)
write.csv(int_items,"data/int_items.csv", row.names=FALSE)
write.csv(all_l,"data/all_l.csv", row.names=FALSE)
write.csv(items_l,"data/items_l.csv", row.names=FALSE)
write.csv(ext_items_l,"data/ext_items_l.csv", row.names=FALSE)
write.csv(int_items_l,"data/int_items_l.csv", row.names=FALSE)
write.csv(ditems_l,"data/ditems_l.csv", row.names=FALSE)
write.csv(dext_items_l,"data/dext_items_l.csv", row.names=FALSE)
write.csv(dint_items_l,"data/dint_items_l.csv", row.names=FALSE)
```

```{r echo=FALSE}
rm(list = ls())
```

