# Dimensionality
This chapter deals with the topic of dimensionality. It begins with a definition of dimensionality and explains how it is important. It outlines how to investigate it and suggests what counts as good evidence. Then it reports our study of the dimensionality of the SHAS.

## About Dimensionality
Dimensionality is the study of the dimensions (also known as "factors", "components", and "latent traits") that the items are designed and assembled to collectively measure. This study tends to take two forms. One is exploration of what dimensions a set of items apparently measure, regardless of what they intended to measure. The other is what evidence exist to confirm that a set of items actually measure the dimension they are intended to measure. 

Dimensionality is important for the reporting of scores. Scores, as summary measures of a set of items, are only meaningful to the extent that they enjoy evidence that they measure the dimension(s) they are designed to measure. Scores, in turn, are important to the extent that they support decisions that have consequences in the real world. Scores are needed to decide who gets to practice medicine, or fly an airplane. It follows that dimensionality is an important part of the validation process for psychological measures [@EmbretsonReise2000].

The methodology commonly used to investigate dimensionality is factor analysis. To assess the dimensionality of these 66 piloted SHAS items we use exploratory factor analysis and confirmatory factor analysis (CFA) and PCAR.

## Exploratory Factor Analysis (EFA)

```{r}
# LOAD PACKAGES
library(lavaan)
library(psych)
library(psychTools)

# LOAD DATA
# all         <- read.csv("data/all.csv")
items       <- read.csv("data/items.csv")
items_l <- read.csv("data/items_l.csv")
ext_items   <- read.csv("data/ext_items.csv")
int_items   <- read.csv("data/int_items.csv")

```

```{r}
# TEST FOR NUMBER OF FACTORS IN DATA USING PARALLEL ANALYSIS AND VERY SIMPLE STRUCTURE (VSS)
fa.parallel(items)
vss(items)
```

```{r}
# PRINCIPAL COMPONENTS ANALYSIS
psych::principal(items)
```

```{r}
# PRINCIPAL AXIS FACTORING
# library(psych)
# fit <- factor.pa(items, nfactors=3, rotation="varimax")
# fit # print results
```

```{r}
# PRINCIPAL COMPONENTS ANALYSIS
# entering raw data and extracting PCs
# from the correlation matrix

fit <- princomp(items_l, cor=TRUE)
summary(fit) # print variance accounted for
loadings(fit) # pc loadings
plot(fit,type="lines") # scree plot
# fit$scores # the principal components
# biplot(fit)
```

```{r}
# DETERMINE NUMBER OF FACTORS TO EXTRACT
library(nFactors)
ev <- eigen(cor(items_l)) # get eigenvalues
ap <- parallel(subject=nrow(items_l),var=ncol(items_l),
   rep=100,cent=.05)
 nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
plotnScree(nS)
```

```{r}
# MAXIMUM LIKELIHOOD FACTOR ANALYSIS
# entering raw data and extracting 3 factors,
# with varimax rotation
fit <- factanal(items_l, 6, rotation="varimax")
print(fit, digits=2, cutoff=.3, sort=TRUE)

# # plot factor 1 by factor 2
# load <- fit$loadings[,1:2]
# plot(load,type="n") # set up plot
# text(load,labels=names(items),cex=.7) # add variable names

```



## Summary
