### Exploratory Factor Analysis

@KochEdstrom2022 conducted Kaiser-Meyer-Olkin (KMO) tests on the 61 items to gauge the shared variance among the items. The KMO index was .99, exceeding the .5 threshold required to proceed. 

```{r kmo}
# KMO FACTOR ANALYSIS
library(psych)
psych::KMO(items2)
```

We further replicated @KochEdstrom2022 conducted factor analysis with Varimax rotation on the 59 items.   

```{r}
# EXPLORATORY FACTOR ANALYSIS
efa <- psych::fa(items2, rotate = "Varimax")
print(efa)
```

[Table of factor loadings]

>Factors with an Eigenvalue of 1 or greater were retained (see all Eigenvalues in Appendix C). This resulted in 6 total factors, explaining 62% of variance cumulatively. Each factor was examined based on the items that loaded most strongly and labeled as followed:
 - Maintaining the System
 - Internal Distress
 - Embracing Violence
 - Controlling Leadership
 - Harmful God-image
 - Gender Discrimination

>While factor titles for 1, 3, 4, and 6 describe the type of potentially abusive external events that victims may experience, the title of factor 2, “Internal Distress,” describes internal states that often result from spiritual abuse. Item 5, “Harmful God-image,” also describes personal feelings relating to God as a villain in a person’s story. Items measuring racial and sexual discrimination and medical care were removed earlier due to low mean scores. Although discrimination based on race or sexual orientation is just as pernicious as gender discrimination, the items’ low mean scores make a mathematical argument for inclusion on the final scale more difficult since it is naturally less common across the entire population than gender discrimination. Inclusion in a critical items section of the final instrument allows for the assessment of discrimination on the basis of race or sexual orientation. Another item, “Being deterred from seeking mental health treatment,” is clinically important and is therefore included as a “Critical Item”.
Individual items were then sorted by factor loadings, with only items loading at .42 or higher (Varimax rotation) per factor being retained. Because the final scale is intended primarily for use in clinical settings, perceived clinical relevance was additionally considered when choosing which items to include on the final scale while being sure to maintain high reliability scores. Factor-specific reliability analyses were performed repeatedly, removing one or two items at a time, to reduce the number of items needed per factor. Descriptive statistics for each factor and the full 27-item scale are below in Table 3. 