```{r eval=FALSE, echo=FALSE}
# INITIAL ITEM ANALYSIS
```

```{r}
# CLASSICAL ITEM STATISTICS
```

```{r}
# POLYCHORIC CORRELATION MATRIX
```

>XXX pairs of items in the polychoric correlation matrix were highly correlated (r > 0.70). After removing X items, these pairs disappeared (several items appeared in multiple pairs). Next, we examined how individual items (or small groups of items) might separate from larger sets of items with a hierarchical item cluster plot [34]. This plot showed final clustering of a pair of items with narrow content and a single item that had poor wording and the lowest cluster loading (i.e., corrected item-total correlation). We deleted these three items and an additional item with redundant wording, leaving XX in the item pool at this stage.

### Principal Components Analysis (PCA)

```{r pca-run}
library(lavaan)
library(psych)
library(psychTools)
all <- read.csv("data/all.csv")

# PRINCIPAL COMPONENTS ANALYSIS
# fit <- princomp(all[,1:66], cor=TRUE)
```

```{r}
# REPORT PCA
# summary(fit) # print variance accounted for
# loadings(fit) # pc loadings
# plot(fit,type="lines") # scree plot
# fit$scores # the principal components
# biplot(fit)


# psych::principal(all[,1:66])

```

### Cluster Analysis

```{r iclust-run}
# CLUSTER ANALYSIS
# ic <- psych::iclust(all[1:66])
# print(ic)
# psych::iclust(r.mat = all[,1:66])
# 
# r.poly <- polychoric(all[,1:66],correct=0) 
# ic.poly <- iclust(r.poly$rho,title="ICLUST using polychoric correlations")
# iclust.diagram(ic.poly)
# print(ic,cut=.3)
```
