## Initial Item Analysis

```{r, include=FALSE}
knitr::opts_chunk$set(comment = "", echo=FALSE, warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)
knitr::opts_knit$set(verbose = FALSE)
```

In Table \@ref(tab:CTT-item-table) we report classical statistics for the 66 SHAS items. We sorted the items by their mean. We flagged six items -- EQ41, EQ46, EQ17, EQ55, EQ60, and EQ27 -- due to extreme values of skew (less than -2 or greater than +2) or kurtosis (less than -7 or greater than +7) and item-total correlations below 0.50. Category response proportions show that most people responded "Never" to these items. Thus these items do not apply to the vast majority of people in this population. We removed them from the item pool. 

```{r CTT-item-stats, echo=FALSE}
# RUN CTT STATS, BUILD ITEM TABLE
rm(list=ls())
source("code/Build-CTT-ItemStats-Table.R")
```

```{r CTT-item-table, echo=FALSE}

# RENDER CTT ITEM STATS TABLE
library(kableExtra)
item.table2 %>%
   kbl(digits=2, align = "ccccccccc", longtable = T, booktabs = T, caption = 'Classical Item Statistics for the SHAS Items') %>%
   kable_styling(latex_options = c("repeat_header"), font_size = 7) %>%
   add_header_above(c(" ", "Category response proportion" = 5, " " = 7)) %>%
   column_spec(1, width="2cm")

# A POSSIBLE BAR PLOT OF ITEM STATISTICS
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=TRUE, col='red')
# barplot(item_stats$nbr.val, main="Complete Cases", horiz=FALSE, col='red', ylim = c(0,8000),names.arg=labs)

```

```{r}
# SAVE NEW ITEM FILE
all <- read.csv("data/all.csv")
items2 <- all[ ,2:66]
items2 <- items2[,-c(5,15,28,33,41,47)]
write.csv(items2,"data/items2.csv")
rm(list=ls())
```
