## IRT Dimensionality Assumptions

```{r, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(comment = "", warning = FALSE, echo = FALSE,  message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)
knitr::opts_knit$set(verbose = FALSE)
```

Following @ReiseEtAl2013, we conducted several analyses:

- inspection of eigenvalues from principal component or factor analysis (e.g., how large is the ratio of the first to second eigenvalue?),
- determining dimensionality through parallel analysis
- compared the fit of a one-factor confirmatory model against benchmarks (e.g., comparative fit index > .90)
- inspected residuals after fitting a unidimensional IRT model

First we confirmed the factor structure using confirmatory factor analysis (CFA). Then we determined the proportion of variance in the observed item responses due to the Rasch model. Finally, we conducted Rasch Principal Component Analysis of Residuals (PCAR) (Smith, 2002).

### Eigenvalues

```{r pca-run, echo=FALSE}
# PRINICIPAL COMPONENTS ANALYSIS
# pca <- principal(items2)
# summary(pca)

# Principal Components Analysis
library(psych)
items2 <- read.csv("data/items2.csv")
items2 <- items2[,-1]
items2 <- items2[complete.cases(items2),]
pca <- princomp(items2, cor=TRUE)
# print(pca)
# summary(pca) # print variance accounted for
# loadings(pca) # pc loadings
plot(pca,type="lines") # scree plot
# pca$scores # the principal components
# biplot(pca)
```

```{r}
# CONFIRMATORY FACTOR ANALYSIS
# items2 <- read.csv("data/items2.csv")
# items2 <- items2[,1]
# library(lavaan)

#one factor three items, default marker method
# m1a  <- ' f  =~ q03 + q04 + q05'
# onefac3items_a <- cfa(m1a, data=dat) 
# summary(onefac3items_a) 

# Model with one common factor
# abuse.factor <- ' #start of model

# latent variable definitions (common factors)
  # wellbeing =~ NA*sense.peace + 
  #           NA*sense.wonder +
  #           NA*sense.gratitude +
  #           NA*sense.purpose
  
# latent variable variances
  # wellbeing ~~ 1*wellbeing

# latent variable covariances
# latent variable means

# manifest variable variances (uniquenesses)
  # sense.peace ~~ sense.peace
  # sense.wonder ~~ sense.wonder
  # sense.gratitude ~~ sense.gratitude
  # sense.purpose ~~ sense.purpose
            
# manifest variable covariances (uniquenesses)

# manifest variable means 
  # sense.peace ~ 1
  # sense.wonder ~ 1
  # sense.gratitude ~ 1
  # sense.purpose ~ 1
 # end of model

# CFA1 <- lavaan(wb_1factor, data=items2, estimator = "MLM") # estimate model
# summary(CFA1, standardized=TRUE, fit.measures=TRUE)

```

**Proportion of variance**. In this analysis, we estimated a Rasch model on dichotomized responses to the 59 SHAS items and saved the residuals of the person parameters. We calculated the variance in the observed item responses and the variance of the residuals, then proportion of variance explained by the Rasch model. This value was .23. Reckase (citation) suggests that the unidimensionality assumption is safely met if the Rasch model explains 20% of the variance in the data. Thus we have some evidence of unidimensionality with which to proceed to further IRT modeling.

```{r uni-variance, echo=FALSE}

# 1- PREP DATA
library(psych)
library(eRm)
items2 <- read.csv("data/items2.csv") # NO MISSING DATA
items2 <- items2[,-1]
# source("code/ApplyItemLabels.R") # apply variable labels
items2 <- items2-1 # subtract 1 from all responses to rescale from 0-4
items2[items2 == 2] <- 0
items2[items2 == 3] <- 1
items2[items2 == 4] <- 1
items2 <- items2[complete.cases(items2), ]

# 2 - ESTIMATE THE MODEL
results <- RM(items2)
# summary(results)

# 3 - GET RESIDUALS
person.locs <- person.parameter(results) # person parameters
model.prob <- pmat(person.locs) # model-predicted probabilities for a correct response for each examinee on each item
responses.without.extremes <- person.locs$X.ex # modified response matrix of no extreme values
resids <- responses.without.extremes - model.prob # residuals = difference between observed responses and model predictions

# 4 - CALCULATE PROPORTION OF VARIANCE FROM RASCH MODEL
## Variance of the observations: VO
observations.vector <- as.vector(responses.without.extremes)
VO <- var(observations.vector)

## Variance of the residuals: VR
residuals.vector <- as.vector(resids)
VR <- var(residuals.vector)

## Raw variance explained by Rasch measures: (VO - VR)/VO
#library(knitr)
#knitr::kable((VO - VR)/VO)
(VO - VR)/VO

```

**Principal Component Analysis of Residuals (PCAR)**. Next we examined the principal components of the correlations among residuals. The premise is that once the Rasch model has been estimated, the correlations among the residuals should be minimal. Linacre () suggests that contrasts with eigenvalues of 2.0 or below can be considered noise. In Figure \@ref(fig:uni-pcar), our PCAR analysis found only one contrasts that rose above that 2.0 threshold.

```{r uni-pcar, fig.cap='Contrasts from PCA of Standardized Residual Correlations', echo=FALSE}
# PCA OF RESIDUAL CORRELATIONS
item.fit <- itemfit(person.locs)
std.resids <- item.fit$st.res
pca <- pca(std.resids, nfactors = ncol(items2), rotate = "none")
contrasts <- c(pca$values[1], pca$values[2], pca$values[3], pca$values[4], pca$values[5])
plot(contrasts, ylab = "Eigenvalues for Contrasts", xlab = "Contrast Number")
```

**Q3**. In our check of the assumption of local independence, we calculated the Q3 statistic (Yen, 1984). This statistic examines the correlations among item residuals, the premise being that the latent trait should account for so much common variance in the item responses that any net correlations among the items should be weak. The Q3 statistic index criteria are that the raw residual correlation between pairs of items should never exceed 0.10 (Marais & Andrich, 2008). Of the 1,711 item pairs, the mean correlation was -0.017 with a standard deviation of 0.035. Thus, most of the residual correlations were very weak.

Figure \@ref(fig:q3-plot) plots the matrix of correlations between item residuals. Each square depicts a correlation. The squares are shaded in grey with the lightest shade indicating the weakest correlations. As hoped, the vast majority of squares are very light, indicating weak correlations among the item residuals.

```{r q3, echo=FALSE}
# Q3 STATISTIC
library(sirt)
# items <- read.csv("data/ditems_l.csv") # grab the data
mod <- sirt::rasch.mml2(items2, progress = FALSE) # estimate the Rasch model
mod.wle <- sirt::wle.rasch(dat=items2, b=mod$item$b) # estimate WLEs
mod.q3 <- sirt::Q3(dat=items2, theta=mod.wle$theta, b=mod$item$b) # estimate Q3
```

```{r q3-plot, fig.cap='Correlations matrix of Item Residuals', echo=FALSE}
# plot Q3 statistics
I <- ncol(items2)
image( 1:I, 1:I, mod.q3$q3.matrix, col=grey( 1 - (0:32)/32),
       xlab="Item", ylab="Item")
# abline(v=c(53,66)) # borders for testlets
# abline(h=c(53,66))
```

```{r, echo=FALSE}
rm(list=ls())
```