
```{r, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(comment = "", warning = FALSE, echo = FALSE,  message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)
knitr::opts_knit$set(verbose = FALSE)
```

### IRT Model Assumptions

IRT models make two assumptions of the data. The first is unidimensionality of the data. This is the assumption that "a single common latent variable accounts for all the common variance among the items" [@ReiseEtAl2013].  The SHAS items were multidimensional by design. The construct of spiritual harm and abuse is not conceptually narrow, and Dan wrote two distinct sets of items -- external events and internal states -- to cover the breadth of this construct. The exploratory factor analyses found a handful of factors in the SHAS items. A unidimensional IRT model can be fit to multidimensional data [@ReiseEtAl2013]. We need evidence of a strong common trait.

The second assumption is local independence [@EmbretsonReise2000]. This means that, once the latent variable has been accounted for, any set of items should not share any meaningful correlation (Edwards, Houts, & Cai, 2018).

#### Unidimensionality

Following @ReiseEtAl2013, we conducted several analyses:

- inspection of eigenvalues from principal component or factor analysis (e.g., how large is the ratio of the first to second eigenvalue?),
- determining dimensionality through parallel analysis
- compared the fit of a one-factor confirmatory model against benchmarks (e.g., comparative fit index > .90)
- inspected residuals after fitting a unidimensional IRT model

First we confirmed the factor structure using confirmatory factor analysis (CFA). Then we determined the proportion of variance in the observed item responses due to the Rasch model. Finally, we conducted Rasch Principal Component Analysis of Residuals (PCAR) (Smith, 2002).

**Factor structure**.

**Proportion of variance**.

```{r uni-variance}
source("code/Unidimensionality-Variance.R")
```

Reckase suggests that the unidimensionality assumption is safely met if the Rasch model explains 20% of the variance in the data. According to this analysis, the Rasch model explains 44% of the variance in the SHAS item data. Thus we have some evidence of unidimensionality with which to proceed to further Rasch modeling.

**Principal Component Analysis of Residuals (PCAR)**

```{r uni-pcar, fig.cap='Contrasts from PCA of Standardized Residual Correlations'}
source("code/Unidimensionality-PCAR.R")
```

The second analysis of PCAR examines the correlations among residuals. The premise is that once the Rasch model has been estimated, the correlations among the residuals should be minimal. Linacre () suggests that contrasts with eigenvalues of 2.0 or below can be considered noise. In Figure \@ref(fig:uni-pcar), our PCAR analysis found three contrasts that rose above that 2.0 threshold.

```{r child='_IRT_LocalIndependence.Rmd'}
```

**Summary**

These analyses offer mixed evidence of assumption satisfaction. The model explains 44% of the variance of the item responses. But we see three contrasts of the standardized residuals above the threshold of 2. In the Q3 plot we see a few strong correlations of item correlations.

```{r}
rm(list=ls())
```