## Unidimensionality and Local Independence

```{r, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(comment = "", warning = FALSE, echo = FALSE,  message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)
knitr::opts_knit$set(verbose = FALSE)
```

```{r}
library(psych)
library(eRm)
items2 <- read.csv("data/items2.csv")
items2 <- items2[,-1]
items2 <- items2[complete.cases(items2),]
```

**Principal Components**. We first analyzed the principal components in the SHAS items. Like factor analysis, component analysis attempts to summarize the common variance in items by identifying a smaller set of latent sources of covariation. Principal components analysis differs from factor analysis by treating all of the common variance among the items as the total variance to be explained [verify this].

```{r pca-run, fig.cap='Principal Components Analysis'}
# PRINICIPAL COMPONENTS ANALYSIS

# pca <- principal(items2)
# summary(pca)

pca <- princomp(items2, cor=TRUE)
# print(pca)
# summary(pca) # print variance accounted for
# loadings(pca) # pc loadings
plot(pca,type="lines", main="") # scree plot
# pca$scores # the principal components
# biplot(pca)

# fa.parallel(items2)
```


As shown in the scree plot, the first component was large, accounting for nearly 30% of the variance. The second, third, and fourth components accounted for 4%, 3%, and 2% of the variance, respectively. The ratio of the eigenvalue of the first (largest) component to that of the second is approximately 13.

```{r}
# CONFIRMATORY FACTOR ANALYSIS
# items2 <- read.csv("data/items2.csv")
# items2 <- items2[,1]
# library(lavaan)

#one factor three items, default marker method
# m1a  <- ' f  =~ q03 + q04 + q05'
# onefac3items_a <- cfa(m1a, data=dat) 
# summary(onefac3items_a) 

# Model with one common factor
# abuse.factor <- ' #start of model

# latent variable definitions (common factors)
  # wellbeing =~ NA*sense.peace + 
  #           NA*sense.wonder +
  #           NA*sense.gratitude +
  #           NA*sense.purpose
  
# latent variable variances
  # wellbeing ~~ 1*wellbeing

# latent variable covariances
# latent variable means

# manifest variable variances (uniquenesses)
  # sense.peace ~~ sense.peace
  # sense.wonder ~~ sense.wonder
  # sense.gratitude ~~ sense.gratitude
  # sense.purpose ~~ sense.purpose
            
# manifest variable covariances (uniquenesses)

# manifest variable means 
  # sense.peace ~ 1
  # sense.wonder ~ 1
  # sense.gratitude ~ 1
  # sense.purpose ~ 1
 # end of model

# CFA1 <- lavaan(wb_1factor, data=items2, estimator = "MLM") # estimate model
# summary(CFA1, standardized=TRUE, fit.measures=TRUE)

```

**Proportion of variance**. We estimated a unidimensional Rasch model on dichotomized responses to the 59 SHAS items and saved the residuals of the person parameters. We calculated the variance in the observed item responses and the variance of the residuals. 

```{r uni-variance}

# 1- PREP DATA
items3 <- items2-1 # subtract 1 from all responses to rescale from 0-4
items3[items3 == 2] <- 0
items3[items3 == 3] <- 1
items3[items3 == 4] <- 1

# 2 - ESTIMATE THE MODEL
results <- RM(items3)
# summary(results)

# 3 - GET RESIDUALS
person.locs <- person.parameter(results) # person parameters
model.prob <- pmat(person.locs) # model-predicted probabilities for a correct response for each examinee on each item
responses.without.extremes <- person.locs$X.ex # modified response matrix of no extreme values
resids <- responses.without.extremes - model.prob # residuals = difference between observed responses and model predictions

# 4 - CALCULATE PROPORTION OF VARIANCE FROM RASCH MODEL
## Variance of the observations: VO
observations.vector <- as.vector(responses.without.extremes)
VO <- var(observations.vector)

## Variance of the residuals: VR
residuals.vector <- as.vector(resids)
VR <- var(residuals.vector)

## Raw variance explained by Rasch measures: (VO - VR)/VO
#library(knitr)
#knitr::kable((VO - VR)/VO)
(VO - VR)/VO

```

The proportion of variance explained by the Rasch model was .23. If the unidimensionalty assumption is safely met if the Rasch model explains 20% of the variance in the data [@Reckase1979], then the SHAS data meet this criterion.

**Principal Component Analysis of Residuals (PCAR)**. We examined the principal components of the correlations among residuals of the Rasch analysis. The premise is that once the Rasch model has been estimated, correlations among the item residuals should be minimal. Linacre () suggests that contrasts with eigenvalues of 2.0 or below can be considered noise. 

```{r uni-pcar, fig.cap='Contrasts from PCA of Standardized Residual Correlations'}
# PCA OF RESIDUAL CORRELATIONS
item.fit <- itemfit(person.locs)
std.resids <- item.fit$st.res
pca <- pca(std.resids, nfactors = ncol(items3), rotate = "none")
contrasts <- c(pca$values[1], pca$values[2], pca$values[3], pca$values[4], pca$values[5])
plot(contrasts, ylab = "Eigenvalues for Contrasts", xlab = "Contrast Number")
```

In Figure \@ref(fig:uni-pcar), our PCAR analysis found only one contrast that rose above that 2.0 threshold.

**Q3**. Finally, we calculated the Q3 statistic (Yen, 1984). This statistic examines the correlations among item residuals, the premise being that the latent trait should account for so much common variance in the item responses that any net correlations among the items should be weak. The Q3 statistic index criteria are that the raw residual correlation between pairs of items should never exceed 0.10 (Marais & Andrich, 2008). Of the 1,711 item pairs, the mean correlation was -0.017 with a standard deviation of 0.035. Thus, most of the residual correlations among items were very weak. 

```{r q3}
# Q3 STATISTIC
library(sirt)
mod <- sirt::rasch.mml2(items3, progress = FALSE) # estimate the Rasch model
mod.wle <- sirt::wle.rasch(dat=items3, b=mod$item$b) # estimate WLEs
mod.q3 <- sirt::Q3(dat=items3, theta=mod.wle$theta, b=mod$item$b) # estimate Q3
```

```{r q3-plot, fig.cap='Matrix of Correlations of Item Residuals'}
# plot Q3 statistics
I <- ncol(items3)
image( 1:I, 1:I, mod.q3$q3.matrix, col=grey( 1 - (0:32)/32),
       xlab="Item", ylab="Item")
# abline(v=c(53,66)) # borders for testlets
# abline(h=c(53,66))
```

Figure \@ref(fig:q3-plot) plots the matrix of correlations between item residuals. Each square depicts a correlation. The squares are shaded in grey with the lightest shade indicating the weakest correlations. As hoped, the vast majority of squares are very light, indicating weak correlations among the item residuals.

```{r}
rm(list=ls())
```