#### Local Independence

To check the assumption of local independence, we used the Q3 statistic (Yen, 1984). The Q3 statistic index criteria are that the raw residual correlation between pairs of items should never exceed 0.10 (Marais & Andrich, 2008).

```{r q3, echo=FALSE}
# Q3 STATISTIC
library(sirt)
items <- read.csv("data/ditems_l.csv") # grab the data
mod <- sirt::rasch.mml2(items) # estimate the Rasch model
mod.wle <- sirt::wle.rasch(dat=items, b=mod$item$b) # estimate WLEs
mod.q3 <- sirt::Q3(dat=items, theta=mod.wle$theta, b=mod$item$b ) # estimate Q3
```

```{r q3-plot, fig.cap='Correlations matrix of Item Residuals'}
# plot Q3 statistics
I <- ncol(items)
image( 1:I, 1:I, mod.q3$q3.matrix, col=gray( 1 - (0:32)/32),
       xlab="Item", ylab="Item")
abline(v=c(53,66)) # borders for testlets
abline(h=c(53,66))
```

>(Some sample language) "Using these criteria, we found XX items in possession of local dependence. The items that had the highest raw residual correlations had negative signs and no positive residual correlation. In other words, the assumption of local independence in this study (was)/(was not) met."

Figure \@ref(fig:q3-plot) is a plot of the matrix of correlations between item residuals.