--- 
title: "Validation of the Spiritual Harm and Abuse Scale with the Rasch Rating Scale Model"
author:
  - Jack Huber, Swedish Medical Center
  - Dan Koch, Northwest University
date: "`r format(Sys.time(), '%B %Y')`"
site: bookdown::bookdown_site
documentclass: article
bibliography: [book.bib, packages.bib]
biblio-style: apalike
url: https://jackhuber.org/SHAS/
description: |
  This is a minimal example of using the bookdown package to write a book.
  The HTML output format for this example is bookdown::gitbook,
  set in the _output.yml file.
link-citations: yes
github-repo: https://github.com/jackbhuber/SHAS.git
---
# Abstract
Spiritual harm and abuse has been a serious problem that has begun to gain the attention of the clinical and psychological research community. Some research has attempted to define spiritual harm and abuse and documented lasting emotional and psychological damage to victims of spiritual abuse. A valid scale of spiritual and religious abuse is needed, and the nascent Spiritual Harm and Abuse Scale (SHAS) has potential to meet this need. @KochEdstrom2022 developed an item pool and conducted an initial validation study of the SHAS which established the multidimensional structure of the item pool. This study builds on @KochEdstrom2022 to further validate the SHAS by using the Rasch Rating Scale Model.

```{r, include=FALSE}

knitr::opts_chunk$set(comment = "", warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)

knitr::opts_knit$set(verbose = FALSE)

# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(),'bookdown','eRm','lavaan','ltm','knitr','lordif','mirt','psych','rmarkdown','sirt','TAM','VIM'), 'packages.bib')
```

<!--chapter:end:index.Rmd-->

# Introduction

Spiritual and religious abuse has been a serious problem which is gaining the attention of the psychological research community. @KochEdstrom2022 define spiritual abuse as "a type of emotional and psychological abuse perpetrated by a religious leader or group and/or with a religious or spiritual component, usually involving coercion or control." Such abuse can range from psychologically controlling behaviors like quoting scripture or invoking divine rationale for mistreatment (Oakley et al., 2018); to diverse emotionally abusive behaviors such as rejecting, isolating, terrorizing, ignoring, corrupting, verbally assaulting, and over-pressuring (Pearl, 1994); to sexual or other physical forms of abuse by a clergy member (Doyle, 2006). Spiritual abuse can also range in specificity from abusive acts by religious leaders to abuse perpetrated by a religious group or person representing a religious group with a religious or spiritual component (Swindle, 2017). Spiritual abuse may also include theologically sanctioned mistreatment of a group of people such LGBTQ+ persons (Foster, Bowland, & Vosler, 2015). 

Spiritual and religious abuse has serious negative effects (Ward, 2011; Swindle, 2017; Oakley & Kinmond, 2014; Stevens et al. 2019). Religious leaders may be on a par with parents in their potential for abusive harm. Further, trauma that takes place in a religious or spiritual setting may deprive victims of religion and spirituality as a coping strategy. There are also concerns related to secondary trauma for clinicians working with R/S abused clients (Gubi & Jacobs, 2009).

In light of these concerns, a valid scale of spiritual harm and abuse is needed. Practitioners who treat clients who may be victims of religious and spiritual abuse need an instrument for clinical use. Clinicians could use such a tool as part of the intake process, to evaluate their client’s score in light of established norms. In addition, researchers need a reliable measurement of spiritual abuse to examine relationships between spiritual abuse and any number of other psychological, health, or other measures.

Drawing from a recent unpublished scale of religious and abuse [@Keller2016] and other scales of abuse (Kira, et al. 2008; Sanders & Becker-Lausen, 1995; Carlson, et al. 2011; Briere & Runtz, 1989), @KochEdstrom2022 developed the Spiritual Harm & Abuse Scale (SHAS). These authors drew from the literature and interviews to find a wide range of experiences and emotional states that could be described as spiritual harm and abuse. From this work they developed a pool of 66 prompts for an online community survey which they fielded in 2021. Their study of the factor structure revealed a handful of factors which they used to distill items to a 27-item scale with six subscales.

In this paper, we build upon the work of @KochEdstrom2022 to further validate the SHAS by using item response theory. Item response theory can enhance clinical assessments in several ways. First, IRT provides interval-level measures which are more appropriate for statistical analysis than ordinal-level sum scores. Second, IRT improves upon sum scores by providing a scale along which both items and persons can be located. Third, measurements derived from IRT are sample-free, meaning it is possible to estimate a person's level of the latent construct free of the distribution of individual items and to estimate an item's difficulty level free from the distribution of people used in the sample. Finally, IRT improves upon single overall reliability coefficients by providing examination of precision across the score continuum.

The goal of this study was thus to use the Rasch Rating Scale Model (an IRT model) to enhance the SHAS in several ways. First, we aimed to express the SHAS as a single interval-level outcome measure of severity of spiritual harm and abuse. Second, we aimed to enrich the diagnostic value of the items and scale by examining severity. Third, we aimed to examine the distribution of item severities to see where the scale provides the most precise, reliable measurement of spiritual abuse. Such a scale also lays the foundation to further develop the SHAS by identifying anchor items that can calibrate with pilot items on future forms. Finally, we aimed to establish a SHAS score that is free of differential item functioning based on race, gender, and other demographic variables.

<!--chapter:end:01-Introduction.Rmd-->

# Methods

## Item Pool Development

@KochEdstrom2022 developed an initial pool of items to measure spiritual harm and abuse. The authors validated the construct of spiritual harm and abuse through literature review (Kvarfordt, 2010; Nobakht & Dale, 2018; Oakley & Kinmond, 2014; Rodríguez-Carballeira, et al. 2015; Swindle, 2017; Ward, 2011, Winell, n.d.) and interviews with survivors of religious and spiritual abuse. These interviews revealed an array of specific examples of potential abuse not described in the published literature such as abuse inflicted by narcissistic persons, financial coercion, developmentally inappropriate children’s teachings, pressure to stay in physically abusive marriages, neglect of needed medical care, sexual discrimination, and shunning/shaming, among others. The authors generated items by drawing from publicly available measures of similar constructs and by writing new items to maximize construct coverage. Each item they adapted to measure either a potentially spiritually abusive event (inflicting emotional abuse by a religious leader or group and/or with a religious or spiritual component, usually involving coercion or control) or a theorized effect of spiritual abuse, based on the reports of self-described survivors of spiritual abuse in the qualitative literature. In developing items, the authors sought to balance construct coverage with alternate wording and efficiency for participants. They wanted to cover as much of the domain as possible in a survey that could be completed in 15 minutes.

## Participants and Procedures

Participants in the initial SHAS validation study were adults who responded to an online survey. To qualify for the survey, participants had to be at least 18 years of age and to have identified themselves as Christian in the past [@KochEdstrom2022].

The second author used snowball sampling to recruit participants to the study. He promoted the survey through various podcast feeds (in audio form during a podcast episode), e-mail lists, Facebook groups, and other online groups associated with progressive Christian podcasts based in the United States. Listeners/readers were encouraged to invite others who satisfied the criteria to take part in the survey. The second author also posted the link on his own social media accounts with a similar encouragement to invite others who satisfied the study criteria. The sampling priority was to reach the largest, most inclusive, community sample. Participants were consented before beginning the survey. The study was approved by the IRB at Northwest University. Data collection began on January 28, 2021 and concluded on February 27, 2021.

## Measures

**Demographics**

Demographic variables included gender, sexual orientation, race, raised in a Christian home, denomination, current religious identity, age, and theological self-description.

**Spiritual Harm and Abuse items**

The final survey instrument included a total of 66 prompts separated into two distinct categories [@KochEdstrom2022]. The first section, External Events, included 52 prompts about potentially abusive experiences. Prompts in this section measured prevalence by asking “How often have you experienced the above in a church or Christian group setting?” and providing participants a Liker scale of five response categories: (1) Never, (2) Once or twice, (3) Sometimes, (4) Often, and (5) All the time. The large number of prompts in this first section were presented in a randomized order so that all prompts would receive responses by approximately the same number of participants even if many participants did not complete all of the prompts in this section.

The second section, Internal States, included 14 prompts about personal feelings resulting from the abusive experiences described in the first section. Prompts in this section measured prevalence by asking “At any point, how often have you experienced the above as a result of negative religious experiences?” Participants were provided the same 5-point Likert scale of response categories as in the first section: (1) Never, (2) Once or twice, (3) Sometimes, (4) Often, and (5) All the time.

## Data Analysis

In their initial validation of the SHAS, @KochEdstrom2022 examined the dimensionality of the 66 items. The authors used exploratory factor analysis (EFA) to determine the factor structure of the items. In these EFA analyses, the items loaded onto six inter-related factors of eigenvalues of at least 1. This enabled the authors to distill the item pool from 66 to 27 items and to identify several meaningful subscales for potential future use by clinicians and researchers.

For this study, we focused on the commonality of the items in order to use item response theory to establish a single interval scale of spiritual abuse. Item response theory models make two assumptions: (1) a single latent trait explains responses to items (unidimensionality) and (2) after controlling for this latent trait, items are weakly correlated (local independence) [@EmbretsonReise2000; @ReiseWaller2009; @ReiseBonifayHaviland2013]. The SHAS items were multidimensional by design. The construct of spiritual harm and abuse is not conceptually narrow, and to cover the breadth of this construct the Koch wrote two distinct item sets. However, one can still appropriately fit a unidimensional IRT model to multidimensional data [@ReiseMooreHaviland2013]. In this study we hypothesized a general factor of spiritual abuse as the primary cause of responses to the full set of 66 SHAS items. We therefore sought evidence of unidimensionality and local independence in order to calibrate the items and participants to a single IRT scale.

We began by examining classical item statistics (skew, kurtosis, item-total correlation) to identify poorly functioning items. Then we conducted several analyses of dimensionality in search of sufficient evidence of unidimensionality and local independence. This included principal components analysis to compare the eigenvalues of the largest and second largest components. We estimated a Rasch model and calculated the proportion of variance in the data it explained. This we followed with a Principal Components Analysis of Residuals (PCAR) to gauge the importance of components in the residuals. We concluded by estimating the Q3 matrix of correlations among item residuals.

We estimated a Rasch Rating Scale model to the SHAS items. We began by examining plots of the category characteristic curves of the items to visualize step parameters. Then we turned to item severity parameters and item fit statistics to inspect the technical quality of the items. We examined an item-person map to understand the fit of the items to the participants. We concluded by assessing the fit of the model to the data.

Finally, we conducted differential item functioning (DIF) analysis of the items to examine item bias by race, gender, and age. [We applied logistic ordinal regression with IRT scoring. We used the Chi-squared likelihood-ratio statistic as the initial DIF detection criteria (alpha < 0.01), and a cut-off of McFadden pseudo R2Δ ≥ 0.02 in model comparisons to determine substantial DIF, a reasonable threshold used in the development of self-reported health outcomes.

We conducted all analyses in **R** [@R-base]. We used the **psych** package [@R-psych] and **eRm** [@R-eRm] for analyses of dimensionality, the **mirt** [@R-mirt] and **TAM** [@R-TAM] packages for IRT analyses, and the **lordif** [@R-lordif] package for DIF analyses.

<!--chapter:end:02-Methods.Rmd-->

# Results

```{r, child = "_ParticipantCharacteristics.Rmd"}
```

```{r, child = "_InitialItemAnalysis.Rmd"}
```

```{r, child = "_IRT_Unidimensionality.Rmd"}
```

```{r, child = "_IRT_RatingScaleModel.Rmd"}
```

```{r, child = "_IRT_DIF.Rmd"}
```

<!--chapter:end:03-Results.Rmd-->

# Discussion

The purpose of this study was to further validate and enhance the Spiritual Harm and Abuse Scale (SHAS) with the use of the Rasch Rating Scale model. The present study improves upon the sample dependency of factor analysis results and contributes new understanding about the psychometric characteristics of the SHAS.

In this study, we examined unidimensionality, Rasch person/item separation reliability, model-data fit, distribution of the variable map based on the locations of persons and items, and individual item qualities of the SHAS. In terms of dimensionality, we found satisfactory evidence to unidimensionality to apply the Rasch Rating Scale model. We found the SHAS highly reliable using Rasch person/item separation reliability statistics, indicating that most of the items are sensitive enough to capture different levels of spiritual abuse and most persons' observed values overlap with the expected values. Similarly, there was close correspondence between the data and the Rasch Rating Scale model except for the outliers in the XXX ends of the continuum. This indicates reasonable model-data fit. Regarding the variable map, our results indicate that the SHAS, overall, captures a range of item severities that reflect the order structure of spiritual abuse. However, we also found a shortage of items at the very high and low ends of the continuum. The SHAS can capture people who have experienced moderate severity of spiritual abuse, but it is not sensitive enough to discriminate among people who have experienced very severe and less severe spiritual abuse. [For example, based on our Rasch Rating scale model analysis, someone with a measure of − 3.00 logits and another with a measure of − 2.00 logits would be more likely to respond to the SHAS in the same way because there is no item on the scale that targets very low levels of spiritual abuse. Hence, we would not be able to detect which person has relatively more or less spiritual abuse in comparison to each other. It would be similarly impossible to distinguish levels of spiritual abuse between a person with a measure of 3.00 logits and another with a measure of 2.00 logits, and decide who has experienced more severe spiritual abuse due to lack of items at the high end of the continuum.] Finally, in terms of item quality, XX of the XX items exhibited good psychometric quality, whereas XX items were misfits. XXX items were redundant because they provided the exact same information with identical pairs of each item, displaying the same contributions to the sensitivity of the SHAS. Hence, they can be removed from the scale to shorten the duration of administration.

Use of the Rasch measurement model in our study provided a powerful analysis of the functioning of the SHAS at the item level and offered in-depth insights on applicability of the scale in wider contexts due to its sample independent nature. Whereas factor analytic approaches such as exploratory factor analysis and confirmatory factor analysis do not provide invariant measurement due to their sample dependency, the Rasch measurement model is independent from the items that form the scale and independent from the sample that the scale is administered to. This feature will increase the applicability of the scale in wider contexts and replicability of the present study in different populations due to the possibility of direct application of the standards developed in one study to data administered in another study (Yan and Mok 2012).

In conclusion, the results of this study provide strong evidence for the validation of the SHAS despite the need for considering the deletion of XX items that are misfits and of XX items that are redundant. The revision or development of new items can also be considered to capture gradations of challenges at the very high and low ends of the continuum to detect people with very high and low levels of spiritual abuse.

## Implications and future directions

This study presents the validation of the SHAS using the Rasch measurement model, for assessing spiritual abuse in adults. The findings have important implications for clinicians and researchers who are concerned about spiritual abuse. The finding that the majority of the SHAS items were of high quality implies that the items provide adequate information to measure moderate levels of spiritual abuse. However, we found the SHAS less able to distinguish people with very high and low levels of spiritual abuse. Hence, for now clinicians and researchers should be cautious when interpreting their results for people who have very high and low spiritual abuse levels.

We echo the same limitations of these data noted in @KochEdstrom2022. The SHAS was developed on a largely white and Protestant/former Protestant population. Possibly major types or sub-themes of spiritual abuse not captured here are more prominent in other religious traditions. This would be a fruitful area of future study. Research into spiritual abuse among the LGBTQ+ community might also yield insights specific to that population. Future research could also distinguish normed scores of this scale for different populations and denominations.

The final screener has the potential to both improve therapy for innumerable clients who have experienced negative religious experiences and to contribute a knowledge base for any future clinical training around working with spiritually abused clients. The scale may also contribute to future research that could be critical in understanding not only the mechanisms and experience of spiritual abuse but also perhaps the enduring power of religion, both to heal and to harm.

<!--chapter:end:04-Discussion.Rmd-->

# Data Availability 

Data will be made available in accordance with xxx. Within one year of publication of manuscripts addressing the aims of the grant, investigators will make a deidentified, anonymized dataset available to the public.

<!--chapter:end:08-DataAvailability.Rmd-->

# Code Availability

Code for the analyses presented in this paper will be made available upon request.

<!--chapter:end:09-CodeAvailability.Rmd-->

# Acknowledgments

This manuscript was not supported by grant funding. The views expressed in this article are those of the authors and do not necessarily reflect the position or policy of the Swedish Medical Center or Northwest University.

<!--chapter:end:11-Acknowledgments.Rmd-->

# Funding

This research was not supported by grant funding.

<!--chapter:end:12-Funding.Rmd-->

# Author Information

## Affiliations

**Department of Pharmacy, Swedish Medical Center, Seattle, USA**

Jack Huber

**Department of Psychology, Northwest University, Kirkland, USA**

Dan Koch


## Corresponding Author

Correspondence to [Jack Huber](mailto:jack.bernard.huber@gmail.com).

<!--chapter:end:13-AuthorInformation.Rmd-->

# Ethics Declarations

## Conflict of Interest

The authors declare no conflict of interest.

<!--chapter:end:14-EthicsDeclaration.Rmd-->

# References {-}

<div id="refs"></div>


<!--chapter:end:15-References.Rmd-->

