In total, 3,222 individuals responded to the survey. Participants who did not opt out and completed at least 50% of the spiritual abuse item pool were considered complete responders and included in the study (N=3,219). Non-responders included those who responded to less than half of the items (N=3).

```{r dems-prep, echo=FALSE}
# GET DEMOGRAPHIC DATA
library(psych)
dems.itms <- as.data.frame(all[,c(67:81)])
names(dems.itms) <- c(
  Q84 = "Age",
  Q85 = "Gender",
  Q86 = "Sexual orientation",
  Q87 = "Race",
  Q88 = "Raised in Christian home",
  Q89 = "Denomination of most abuse",
  Q90 = "Racial makeup of abusive community",
  Q91 = "Racial makeup of abusive group leadership",
  Q92 = "Gender makeup of abusive group leadership",
  Q95 = "Involvement in abusive group",
  Q93 = "Current religious identification",
  Q97 = "Current theological identification",
  Q98 = "Current view of Bible",
  Generation = "Generation",
  Podcast_Usage = "Podcast Use"
)
library(expss)
dems.itms = expss::apply_labels(dems.itms,
                            Gender = c("Cisgender Male"=1, "Cisgender Female"=2, "Transgender Male"=3, "Transgender Female"=4, "Other"=5, "Prefer not to say"=6),
                            `Sexual orientation` = c("Straight"=1, "Gay/Lesbian"=2, "Bisexual"=3, "Asexual"=4, "Queer"=5, "Don\'t know"=6),
                            Race= c("Ethnically of Hispanic/Latino Origin"=1, "Black/African American"=2, "White/European American"=3, "Asian"=4, "American Indian/Alaska Native"=5, "Native Hawaiian/Pacific Islander"=6, "Other"=7),
                            `Raised in Christian home` = c("Yes"=1, "No"=2),
                            `Denomination of most abuse` = c("Catholic"=1, "Charismatic/Pentecostal"=2, "Baptist"=3, "Mainline"=4, "Non-denominational"=5, "Orthodox Christian"=6, "Mormon"=7, "Christian other than above"=8, "Evangelistic/Missions organization"=9, "College Christian organization"=10, "Christian Youth Organization"=11, "Christian media company"=12, "Christian school or university"=13, "Christian political/activist group"=14, "Christian aid/social services group"=15, "Local Bible study or small group"=16, "Christian group other than the above"=17),
                            `Current religious identification` = c("Protestant"=1, "Catholic"=2, "Mormon"=3, "Orthodox Christian"=4, "Other religious tradition"=5, "Atheist"=6, "Agnostic"=7, "Nothing in particular"=8)
)
write.csv(dems.itms,"data/dems.itms.csv")
```

```{r dems-describe, echo=FALSE}
# DESCRIPTIVE STATISTICS FOR DEMS
dems.tbl <- as.data.frame(describe(dems.itms))
dems.tbl <- subset(dems.tbl[,-1])
dems.tbl <- round(dems.tbl,3)
# dems <- subset(dems[-c(16:18),])
```

```{r dems-table, echo=FALSE}
# source("code/DemsFrequencies.R")
```

Table \@ref(tab:dems-table) shows the demographic and clinical characteristics of the sample used for item analysis. Participants were mostly XXX and XXX but with sizable minority participation (i.e., XX% XXX, XX% XXX).