### Rating Scale Model (RSM)
The first IRT model is the Rating Scale Model (RSM) (citation). The model is an extension of the Rasch model (designed for dichotomous items) for use with polytomous items based on rating scales such as Likert scales of agreement [@EmbretsonReise2000] and estimates the probability of choosing from among several response options given an examinee's location on the latent trait (level of lifetime exposure to spiritual abuse). More specifically, the increase in difficulty involved in choosing from between two response categories is called a "step parameter". The RSM assumes that all items use the same rating scale. As a Rasch model, it assumes that all items discriminate equally well and thus fixes the discrimination parameter to 1. The model also fixes the step parameters of all the items on the premise that the response categories are the same for all items. What the RSM frees to vary are the item difficulty parameters. Thus the focus of the RSM is the differences in item difficulties. If all of these assumptions are true -- that the items and their rating scales differentiate examinees equally well and differ only in overall difficulty -- then the observed pattern of item responses will closely match the pattern of items responses expected by the RSM and the model will fit.

```{r rsm-run}
# ESTIMATE RSM MODEL
library(stats4)
library(lattice)
library(mirt)
items <- read.csv("data/items_l.csv")
# source("code/ApplyItemLabels.R")
model.rsm <- 'SPIRITUAL ABUSE RATING SCALE MODEL = 1-66'
results.rsm <- mirt(data=items, model=model.rsm, itemtype="rsm", verbose=FALSE) # estimate model

# GET ITEM PARAMETERS
coef.rsm <- coef(results.rsm, IRTpars=TRUE, simplify=TRUE) # save coefficients
items.rsm <- as.data.frame(coef.rsm$items) # item parameters as data frame
items.rsm2 <- items.rsm[order(items.rsm$c),]
```

```{r rsm-items, fig.cap='Rating Scale Model Item Parameters for Spiritual Harm and Abuse Items'}
knitr::kable(items.rsm2, digits = 2, "simple")
```

**Item parameters**. Table \@ref(tab:rsm-items) reports the item parameters estimated by the RSM for all 66 SHAS items. The first column shows the discrimination parameter, fixed to “1” for all items on the premise that the items discriminate equally well. The subsequent columns (b1 to b4) report the estimated thresholds (or step parameters). Because the SHAS items all used the same 5-point Likert scale, and because the RSM assumes this scale discriminated examinees equally well, these four threshold parameters are the same for each item. 

What does vary between items, in this model, is the difficulty parameter, shown in the "c" column, which represents the location of the item on the continuum of the latent trait: Does the item represent the "most" spiritual abuse? The least? Or somewhere in the middle?

```{r rsm-item-diff, fig.cap='Distribution of Item Difficulty'}
hist(items.rsm2$c, breaks=50, main = NULL, col="red", xlab="Lifetime Exposure to Spiritual Abuse", xlim = c(-3,3))

# stripchart(items.rsm2$c, xlab = 'Lifetime Exposure to Spiritual Abuse', xlim = c(-3,3), col = 'red', pch = 1, method = 'stack')
```

Figure \@ref(fig:rsm-item-diff) presents a histogram of the item difficult parameters. They are normally distributed but there are extremes.

**Item plots**. Item plots help show this information visually. The following figures are plots of the _option characteristic curves_ (also called category characteristic curves) of the items.

```{r rsm-a}
# RSM ITEM PLOTS
library(directlabels)
a <- plot(results.rsm, type = 'trace', which.items = c(1,2,3,4,5,6),
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4))
direct.label(a, 'top.points')
```

In these plots, the x axis is the scale of the latent trait (lifetime exposure to spiritual abuse) expressed in logits ranging from -6 to +6. The y axis is the probability of selecting a given response (such as "Always") at each level of latent trait (overall lifetime exposure to spiritual abuse). The blue line is the probability of selecting "Never" for examinees ranging from low to high lifetime spiritual abuse. The reversal is that examinees just below zero are more likely to select "Never" than "Once or twice." In addition, examinees with above-average lifetime exposure to spiritual abuse were more likely to select "Always" than "Often". The item parameter statistics shown Table \@ref(tab:pcm-items) indicate that this issue applies to 43 items. These items might contribute more reliable information if these response categories were collapsed.

```{r rsm-item-info}
plot(results.rsm, type='infotrace', which.item = c(1,2,3,4,5,6), facet_items=T, as.table = TRUE, auto.key=list(points=F, lines=T, columns=1, space = 'right', cex = .8), theta_lim = c(-3, 3), main="")
```

>In polytomous models, the amount of information an item contributes depends on its slope parameter—–the larger the parameter, the more information the item provides. Further, the farther apart the location parameters (b1, b2, b3, b4), the more information the item provides. Typically, an optimally informative polytomous item will have a large location and broad category coverage (as indicated by location parameters) over theta.

>Information functions are best illustrated by the item information curves for each item as displayed above. These curves show that item information is not a static quantity, rather, it is conditional on levels of theta. The relationship between slopes and information is illustrated here. Item 3 had the lowest slope and is, therefore, the least informative item.On the other hand, Item 9 had the highest slope and provides the highest amount of statistical information. Items tended to provide the most information between -2.5 to + 1 theta range. The “wavy” form of the curves reflects the fact that item information is a composite of category information, that is, each category has an information function which is then combined to form the item information function.

```{r rsm-itmplot-emp}
itemfit(results.rsm, empirical.plot = 2)
```

**Item fit statistics**. Fit statistics gauge the extent to which the observed pattern of item responses fit the IRT model (Rasch, partial credit, etc.). They do this by means of chi-square statistics which examine the cumulative difference the observed pattern of item responses and the pattern of item respones that the model would expect. Two fit statistics commonly used in IRT modesl are the infit mean square (MNSQ) and the outfit mean square (MNSQ) [@BondFox2015]. The infit statistic places greater emphasis on unexpected responses that are close to the examinees and item location. The outfit is sensitive to unexpected responses that are far from the location. The expected value of infit or outfit for each item is 1.0, with a range of acceptable values ranging from 0.5 to 1.5. Values outside these boundaries indicate a lack of fit between items and the model.

```{r rsm-item-fits}
# ITEM FIT
ifit <- mirt::itemfit(results.rsm, na.rm=TRUE, 'infit')
i <- as.data.frame(itemfit(results.rsm, na.rm = TRUE, 'infit'))
misfits <- subset(i, outfit > 1.5 | outfit < 0.5 | infit > 1.5 | infit < 0.5)
```

```{r rsm-item-fit}
knitr::kable(misfits, digits = 2, caption = "Rating Scale Model (RSM) - Misfitting Items", row.names = FALSE, "simple")
```

All but six of the 66 items had infit and outfit statistics within the acceptable range from 0.5 to 1.5. Statistics for these six misfitting items are presented here in Table \@ref(tab:rsm-item-fit). The large outfit values mean the observed patterns of responses to these items differed substantially from the patterns of responses expected by the PCM. These items capture the experiences of women, people of color, and LGBTQ persons.

```{r}
# infit <- fit$itemfit$Infit
# upper_bound <- rep(x = 1.33, times =15) # this repeats 1.33 fifteen times
# lower_bound <- rep(x = .75, times = 15) 

# running fitgraph
# fitgraph(fitEst = infit, fitLB = lower_bound, fitUB = upper_bound, itemLabels = names(hls))
```

**Model fit**. To assess the overall fit of the model to the item response data we use two statistics. The is the Root Mean Square Error of Approximation (RMSEA). The suggested cutoff for RMSEA is .06. The second is the CFI. The suggested cutoff for this is .95.

```{r rsm-fit, fig.cap='Rating Scale Model Fit - Overall Model Fit'}
# MODEL FIT
# knitr::kable(M2(results.rsm, type = "C2", na.rm=TRUE, calcNULL = FALSE))
knitr::kable(M2(results.rsm, type = "C2"), digits = 2, "simple")
```

Table \@ref(tab:rsm-fit) reports the model fit statistics for the SHAS items. The obtained RMSEA value = .092 (95% CI[.092, .093]) exceeds the cutoff of .06. The CFI = .953 meets the recommended .95 threshold. Together these statistics offer weak evidence that the RSM fits the SHAS items.

```{r}
rm(list = ls())
```