
### IRT Model Assumptions
IRT makes two assumptions of the data. The first is unidimensionality of the latent trait. Is the trait one dimension? This is important for the SHAS because the questionnaire explicitly distinguished between items measuring external events and internal states. To check this assumption we first confirm the factor structure using confirmatory factor analysis (CFA) and Rasch Principal Component Analysis of Residuals (PCAR; Smith, 2002). The second assumption is local independence [@EmbretsonReise2000]. This means that any set of items should not share any meaningful correlation once the latent variable has been accounted for (Edwards, Houts, & Cai, 2018).

## Unidimensionality
First, we will calculate the proportion of variance in responses that can be attributed to item and person locations on the primary latent variable, as estimated using the Rasch model. To do this, we need to find the variance associated with the responses and the variance associated with residuals so that we can compare these values. To do this, we need to calculate person parameters by applying the person.parameter() function to the model results object:

```{r}
# GET PERSON PARAMETERS FROM THE RSM MODEL
responses.without.extremes
```

Then, we can apply the pmat() function to the student locations object to calculate the model-predicted probabilities for a correct response for each student on each of the transitive reasoning tasks:

```{r}
# GET MODEL PREDICTED PROBABILITIES
model.prob <- 

```


Now we can calculate residuals as the difference between the observed responses and model-predictions:

```{r}
resids <- responses.without.extremes - model.prob
```

Following Linacre (2003), we can calculate the proportion of variance in the responses associated with the Rasch model estimates using the following values:

Variance of the observations: VO
Variance of the residuals: VR
Raw variance explained by Rasch measures: (VO - VR)/VO

```{r}
residuals(results.rsm, type='Q3')
```


## Local independence
To check the assumption of local independence, we used the Q3 statistic (Yen, 1984). The Q3 statistic index criteria are that the raw residual correlation between pairs of items should never exceed 0.10 (Marais & Andrich, 2008).

```{r q3, eval=FALSE, include = FALSE}

```

Using these criteria, we found XX items in possession of local dependence. The items that had the highest raw residual correlations had negative signs and no positive residual correlation. In other words, the assumption of local independence in this study (was)/(was not) met.