### IRT Model Assumptions

IRT makes two assumptions of the data. The first is unidimensionality of the latent trait. Is the trait one dimension? This is important for the SHAS because the questionnaire explicitly distinguished between items measuring external events and internal states. The second assumption is local independence [@EmbretsonReise2000]. This means that, once the latent variable has been accounted for, any set of items should not share any meaningful correlation (Edwards, Houts, & Cai, 2018).

#### Unidimensionality

To verify the assumption of unidimensionality we conducted three analyses. First we confirmed the factor structure using confirmatory factor analysis (CFA). Then we determined the proportion of variance in the observed item responses due to the Rasch model. Finally, we conducted Rasch Principal Component Analysis of Residuals (PCAR) (Smith, 2002).

**Factor structure**.

**Proportion of variance**.

```{r uni-variance}
source("code/Unidimensionality-Variance.R")
```

Reckase suggests that the unidimensionality assumption is safely met if the Rasch model explains 20% of the variance in the data. According to this analysis, the Rasch model explains 44% of the variance in the SHAS item data. Thus we have some evidence of unidimensionality with which to proceed to further Rasch modeling.

**Principal Component Analysis of Residuals (PCAR)**

```{r uni-pcar, fig.cap='Contrasts from PCA of Standardized Residual Correlations'}
source("code/Unidimensionality-PCAR.R")
```

The second analysis of PCAR examines the correlations among residuals. The premise is that once the Rasch model has been estimated, the correlations among the residuals should be minimal. Linacre () suggests that contrasts with eigenvalues of 2.0 or below can be considered noise. In Figure \@ref(fig:uni-pcar), our PCAR analysis found three contrasts that rose above that 2.0 threshold.

```{r child='_LocalIndependence.Rmd'}
```

**Summary**

These analyses offer mixed evidence of assumption satisfaction. The model explains 44% of the variance of the item responses. But we see three contrasts of the standardized residuals above the threshold of 2. In the Q3 plot we see a few strong correlations of item correlations.